---
title: "Untitled"
author: "Laura_Hakkinen"
date: "2024-09-24"
output: html_document
---

# 1. Upload plant diversity data

```{r}
library("cowplot")
library("dplyr")
library("ggplot2")
library("vegan")
library("tibble")

plant.div <- read.csv2("C:\\Users\\lehakkin\\PhD\\Fungal_ITS_soil_vertical_profile\\Plant_diversity_Yoni_2011_2012.csv", dec = ".")

plant.div$Year <- as.factor(plant.div$Year)
```


# 2. Non-crop vegetation cover (%)

## 2.1. Year 2011

Let's only include organic and conventional in the non-crop coverage analysis as meadow does not have a crop

```{r}
## first lets test if variances are equal so we know if use ANOVA or Kruskal test:
library(car)

df <- subset(plant.div, Year==2011)
df <- subset(df, Treatment!="meadow")

# Using leveneTest()
result = leveneTest(non_crop_cover ~ Treatment, df)
# print the result
print(result)

# high p-value means that variances are similar and that we can do ANOVA

# Shapiro-Wilk’s method is widely recommended for normality test and it provides better power than K-S. It is based on the correlation between the data and the corresponding normal scores.

shapiro.test(df$non_crop_cover)

# high p value means that data is normally distributed

res.aov <- aov(non_crop_cover ~ Treatment, data = df)
summary(res.aov)

# As the ANOVA test is significant, we can compute Tukey HSD
TukeyHSD(res.aov)
```


Non-crop vegetation cover in 2011 is significantly different between conventional and organic


## 2.2. Year 2012

```{r}
df <- subset(plant.div, Year==2012)
df <- subset(df, Treatment!="meadow")

# Using leveneTest()
result = leveneTest(non_crop_cover ~ Treatment, df)
# print the result
print(result)

# high p-value means that variances are similar and that we can do ANOVA

# Shapiro-Wilk’s method is widely recommended for normality test and it provides better power than K-S. It is based on the correlation between the data and the corresponding normal scores.

shapiro.test(df$non_crop_cover)

# low p value means that data is not normally distributed, but let's still do ANOVA

# analysis of variance takes the same form as the usual models you'd see in R
# response ~ dependent, data = environmental grouping

res.aov <- aov(non_crop_cover ~ Treatment, data = df)
summary(res.aov)

# As the ANOVA test is significant, we can compute Tukey HSD
TukeyHSD(res.aov)
```

Non-crop vegetation cover in 2012 is significantly different between conventional and organic

## 2.3. Plot

```{r}
pal <- c(meadow = "#fbc02d", organic = "#8a8a8a", conventional =  "#b71c1c")

df <- subset(plant.div, Treatment!="meadow")
df <- droplevels(df)

df$Treatment <- factor(df$Treatment, levels = c("meadow", "organic", "conventional"))

plot_non_crop_cover <- ggplot(df, aes(x = Treatment, y = non_crop_cover, fill = Treatment)) + facet_grid (cols = vars(Year)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text.y = element_text(size = 12, color = "gray25"),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = NULL,
       y = "Non-crop vegetation cover")
plot_non_crop_cover

```

## 2.4. Add CLD letters

### 2.4.1. Year 2011

```{r}
# subset the Year
year11 <- plant.div[plant.div$Year == "2011",]

year11 <- subset(year11, Treatment!="meadow")
year11 <- droplevels(year11)

year11$Treatment <- factor(year11$Treatment, levels = c("meadow", "organic", "conventional"))

# Plot subset
plot_non_crop_cover11 <- ggplot(year11, aes(x = Treatment, y = non_crop_cover, fill = Treatment)) + facet_grid (cols = vars(Year)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text.y = element_text(size = 12, color = "gray25"),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = NULL,
       y = "Non-crop vegetation cover")

# Annotate this plot
plot_non_crop_cover11 <- plot_non_crop_cover11 +
  geom_text(x = 1.5,  y = 94, 
           label = "**", 
           colour = "black") +
  geom_segment(x = 0.8, xend = 0.8, 
           y = 92, yend = 93,
           colour = "black") +
  geom_segment(x = 2.2, xend = 2.2, 
           y = 92, yend = 93,
           colour = "black") +
  geom_segment(x = 0.8, xend = 2.2, 
           y = 93, yend = 93,
           colour = "black") + ylim(0, 98)

plot_non_crop_cover11

```

### 2.4.2. Year 2012

```{r}
# subset the Year
year12 <- plant.div[plant.div$Year == "2012",]

year12 <- subset(year12, Treatment!="meadow")
year12 <- droplevels(year12)

# Plot subset
plot_non_crop_cover12 <- ggplot(year12, aes(x = Treatment, y = non_crop_cover, fill = Treatment)) + facet_grid (cols = vars(Year)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text.y = element_text(size = 12, color = "gray25"),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = NULL,
       y = "Non-crop vegetation cover")


# Annotate this plot
plot_non_crop_cover12 <- plot_non_crop_cover12 +
  geom_text(x = 1.5,  y = 94, 
           label = "*", 
           colour = "black") +
  geom_segment(x = 0.8, xend = 0.8, 
           y = 92, yend = 93,
           colour = "black") +
  geom_segment(x = 2.2, xend = 2.2, 
           y = 92, yend = 93,
           colour = "black") +
  geom_segment(x = 0.8, xend = 2.2, 
           y = 93, yend = 93,
           colour = "black") + ylim(0, 98)

plot_non_crop_cover12
```


```{r}
library(ggpubr)

plot_non_crop_cover <- ggarrange(plot_non_crop_cover11, plot_non_crop_cover12 + rremove("ylab") + rremove("y.text") + rremove("y.axis") + rremove("y.ticks"),
                     ncol = 2, nrow = 1, widths = c(1,0.8))

plot_non_crop_cover

```


# 3. Richness

## 3.1. year 2011

```{r}

df <- subset(plant.div, Year==2011)

# Using leveneTest()
result = leveneTest(richness ~ Treatment, df)
# print the result
print(result)

# high p-value means that variances are similar and that we can do ANOVA

# Shapiro-Wilk’s method is widely recommended for normality test and it provides better power than K-S. It is based on the correlation between the data and the corresponding normal scores.

shapiro.test(df$richness)

# high p value means that data is normally distributed

# analysis of variance takes the same form as the usual models you'd see in R
# response ~ dependent, data = environmental grouping

res.aov <- aov(richness ~ Treatment, data = df)
summary(res.aov)

# As the ANOVA test is significant, we can compute Tukey HSD
TukeyHSD(res.aov)
```


Plant richness in 2011 does not significantly differ between any treatments

## 3.2. Year 2012

```{r}

df <- subset(plant.div, Year==2012)

# Using leveneTest()
result = leveneTest(richness ~ Treatment, df)
# print the result
print(result)

# high p-value means that variances are similar and that we can do ANOVA

# Shapiro-Wilk’s method is widely recommended for normality test and it provides better power than K-S. It is based on the correlation between the data and the corresponding normal scores.

shapiro.test(df$richness)

# high p value means that data is normally distributed

# analysis of variance takes the same form as the usual models you'd see in R
# response ~ dependent, data = environmental grouping

res.aov <- aov(richness ~ Treatment, data = df)
summary(res.aov)

# As the ANOVA test is significant, we can compute Tukey HSD
TukeyHSD(res.aov)
```


Plant richness in 2012 does significantly differs between conventional and organic only

## 3.3 Plot

```{r}
pal <- c(meadow = "#fbc02d", organic = "#8a8a8a", conventional =  "#b71c1c")

plant.div$Treatment <- factor(plant.div$Treatment, levels = c("meadow", "organic", "conventional"))

plot_rich <- ggplot(plant.div, aes(x = Treatment, y = richness, fill = Treatment)) + facet_grid (cols = vars(Year)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) + theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text.y = element_text(size = 12, color = "gray25"),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = NULL,
       y = "Plant richness")

plot_rich
```

## 3.4. Add CLD letters

### 3.4.1. Year 2011

```{r}
# subset the Year
year11 <- plant.div[plant.div$Year == "2011",]

# Plot subset
plot_rich11 <- ggplot(year11, aes(x = Treatment, y = richness, fill = Treatment)) + facet_grid (cols = vars(Year)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text.y = element_text(size = 12, color = "gray25"),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = NULL,
       y = "Plant richness")

# Annotate this plot
plot_rich11 <- plot_rich11 +
  geom_text(x = 1,  y = 34, 
           label = "(ns.)")  + ylim(10, 35)
plot_rich11
```

### 3.4.2. Year 2012

```{r}
# subset the Year
year12 <- plant.div[plant.div$Year == "2012",]

# Plot subset
plot_rich12 <- ggplot(year12, aes(x = Treatment, y = richness, fill = Treatment)) + facet_grid (cols = vars(Year)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text.y = element_text(size = 12, color = "gray25"),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = NULL,
       y = "Plant richness")


# Annotate this plot
plot_rich12 <- plot_rich12 +
  geom_text(x = 2.5,  y = 34, 
           label = "*", 
           colour = "black") +
  geom_segment(x = 1.8, xend = 1.8, 
           y = 32, yend = 33,
           colour = "black") +
  geom_segment(x = 3.2, xend = 3.2, 
           y = 32, yend = 33,
           colour = "black") +
  geom_segment(x = 1.8, xend = 3.2, 
           y = 33, yend = 33,
           colour = "black") + ylim(10, 35)

plot_rich12
```


### 3.4.3. Combine figures

```{r}
library(ggpubr)

plot_rich <- ggarrange(plot_rich11, plot_rich12 + rremove("ylab") + rremove("y.text") + rremove("y.axis") + rremove("y.ticks"),
                     ncol = 2, nrow = 1, widths = c(1,0.75))

plot_rich

```


# 4. Shannon diversity

## 4.1. Year 2011

```{r}

df <- subset(plant.div, Year==2011)

# Using leveneTest()
result = leveneTest(Shannon ~ Treatment, df)
# print the result
print(result)

# high p-value means that variances are similar and that we can do ANOVA

# Shapiro-Wilk’s method is widely recommended for normality test and it provides better power than K-S. It is based on the correlation between the data and the corresponding normal scores.

shapiro.test(df$Shannon)

# high p value means that data is normally distributed

# analysis of variance takes the same form as the usual models you'd see in R
# response ~ dependent, data = environmental grouping

res.aov <- aov(Shannon ~ Treatment, data = df)
summary(res.aov)

# As the ANOVA test is significant, we can compute Tukey HSD
TukeyHSD(res.aov)
```

Shannon in 2011 does not significantly differ between any treatments


## 4.2. Year 2012

```{r}

df <- subset(plant.div, Year==2012)

# Using leveneTest()
result = leveneTest(Shannon ~ Treatment, df)
# print the result
print(result)

# high p-value means that variances are similar and that we can do ANOVA

# Shapiro-Wilk’s method is widely recommended for normality test and it provides better power than K-S. It is based on the correlation between the data and the corresponding normal scores.

shapiro.test(df$Shannon)

# high p value means that data is normally distributed

# analysis of variance takes the same form as the usual models you'd see in R
# response ~ dependent, data = environmental grouping

res.aov <- aov(Shannon ~ Treatment, data = df)
summary(res.aov)

# As the ANOVA test is significant, we can compute Tukey HSD
TukeyHSD(res.aov)
```

Shannon in 2012 does not significantly differ between any treatments

## 4.3. Plot

```{r}
pal <- c(meadow = "#fbc02d", organic = "#8a8a8a", conventional =  "#b71c1c")

plant.div$Treatment <- factor(plant.div$Treatment, levels = c("meadow", "organic", "conventional"))

plot_Shannon <- ggplot(plant.div, aes(x = Treatment, y = Shannon, fill = Treatment)) + facet_grid (cols = vars(Year)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text.y = element_text(size = 12, color = "gray25"),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = NULL,
       y = "Plant Shannon diversity")

plot_Shannon
```

## 4.4. Add CLD letters

### 4.4.1. Year 2011

```{r}
# subset the Year
year11 <- plant.div[plant.div$Year == "2011",]

# Plot subset
plot_Shannon11 <- ggplot(year11, aes(x = Treatment, y = Shannon, fill = Treatment)) + facet_grid (cols = vars(Year)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text.y = element_text(size = 12, color = "gray25"),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = NULL,
       y = "Plant Shannon diversity")

# Annotate this plot
plot_Shannon11 <- plot_Shannon11 +
  geom_text(x = 1,  y = 2.9, 
           label = "(ns.)")  + ylim(1.5, 3)
plot_Shannon11
```

### 4.4.2. Year 2012

```{r}
# subset the Year
year12 <- plant.div[plant.div$Year == "2012",]

# Plot subset
plot_Shannon12 <- ggplot(year12, aes(x = Treatment, y = Shannon, fill = Treatment)) + facet_grid (cols = vars(Year)) +
  geom_boxplot() +
  scale_fill_manual(values = pal) +
  theme(legend.position = "right",
        legend.title = element_blank(),
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text.y = element_text(size = 12, color = "gray25"),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = NULL,
       y = "Plant Shannon diversity")



# Annotate this plot
plot_Shannon12 <- plot_Shannon12 +
  geom_text(x = 1,  y = 2.9, 
           label = "(ns.)")  + ylim(1.5, 3)
plot_Shannon12
```

### 4.4.3. Combine figures

```{r}
library(ggpubr)

plot_Shannon <- ggarrange(plot_Shannon11, plot_Shannon12 + rremove("ylab") + rremove("y.text") + rremove("y.axis") + rremove("y.ticks"),
                     ncol = 2, nrow = 1, widths = c(1,1.63))

plot_Shannon

```






# 5. Combine the figures

## 5.1. With Non-crop vegetation cover

```{r fig.dim = c(14, 6)}
library(ggpubr)

fig <- ggarrange(plot_non_crop_cover, plot_rich, plot_Shannon,
                     ncol = 3, nrow = 1, widths = c(3,4,6))

fig

```

## 5.2. Without Non-crop vegetation cover

```{r fig.dim = c(14, 6)}
fig <- ggarrange(plot_rich, plot_Shannon,
                     ncol = 3, nrow = 1, widths = c(2, 3.1))

fig

```
