---
title: "Untitled"
author: "Laura_Hakkinen"
date: "2024-01-10"
output: html_document
---

# RESULTS STEPS 7:FUNGuild testing

Note! While doing tests, always check the homogeneity of variance result and decide based on that which later test result to use (Anova and Tukey or Kruskal and Wilcoxon)

```{r}
library('phyloseq')
library("cowplot")
library("dplyr")
library("ggplot2")
library("vegan")
library("microbiome")
library("tibble")
library(stringr)
library(reshape2)
library(tidyr)

setwd('C:\\Users\\lehakkin\\PhD\\Fungal_ITS_soil_vertical_profile')

load(file = 'ps_FG_with_NAs')#ps_FG
ps_FG
```


# 1. Test Plant Pathogen

Note:

- Ectomycorrhizal           =   guilds containing "Ectomycorrhizal" from trophic mode Symbiotroph only, NOTE! thiis is same as pure Ectomycorrhizal!!
- Arbuscular Mycorrhizal    =   all guilds containing "Arbuscular Mycorrhizal" from trophic mode Symbiotroph (nor AMF in other TMs)
- Endophyte                 =   Pure endophytes from trophic mode Symbiotroph only
- Plant Pathogen            =   Pure Plant Pathogens from trophic mode Pathotroph only


```{r}

# remove species

z <- as.data.frame(tax_table(ps_FG))

tax <- z[, 1:9]

# rename FUNGuild to species
colnames(tax)[3] <- "species"
tax <- tax %>% as.matrix()

# reassign to phyloseq
x <- ps_FG
tax_table(x) <- tax_table(tax)

x <- aggregate_rare(x, level = "species", detection = 0, prevalence = 0)
x               
# 12 taxa and 140 samples


# remove "Unknown"
#allTaxa = taxa_names(x)
#badTaxa = c("Unknown")
#myTaxa <- allTaxa[!(allTaxa %in% badTaxa)]
#x <- prune_taxa(myTaxa, x)
#x
# 11 taxa and 140 samples

x_RA <- transform(x, 'compositional')

```


```{r}

#create data table
FG_df <-  psmelt(x_RA)
```



```{r}

unique(FG_df$OTU)


taxa <- "Plant Pathogen"

# Construct a data.frame with the selected
# taxonomic group and grouping
# relative
df <- filter(FG_df, OTU == taxa)

library(car)
# Using leveneTest()
result = leveneTest(Abundance ~ sample_type, df)
# print the result
print(result)

```

not homogenous


```{r}
# perform the Kruskal test
kruskal.test(Abundance ~ sample_type, data = df)

pairwise.wilcox.test(df$Abundance, df$sample_type,
                     p.adjust.method = "BH")

x <- df %>%
  group_by(sample_type) %>%
  summarise(mean = mean(Abundance, na.rm = TRUE), se = (sd(Abundance, na.rm = TRUE)/sqrt(length((Abundance)))))
x

```



# 2. Test Endophyte

```{r}
unique(FG_df$OTU)


taxa <- "Endophyte"

# Construct a data.frame with the selected
# taxonomic group and grouping
# relative
df <- filter(FG_df, OTU == taxa)

# Using leveneTest()
result = leveneTest(Abundance ~ sample_type, df)
# print the result
print(result)


```


```{r}
# perform the Kruskal test
kruskal.test(Abundance ~ sample_type, data = df)

pairwise.wilcox.test(df$Abundance, df$sample_type,
                     p.adjust.method = "BH")

x <- df %>%
  group_by(sample_type) %>%
  summarise(mean = mean(Abundance, na.rm = TRUE), se = (sd(Abundance, na.rm = TRUE)/sqrt(length((Abundance)))))
x
```


# 3. Test Trophic modes

```{r}
# remove species

z <- as.data.frame(tax_table(ps_FG))

tax <- z[, 1:9]

# rename trophicmode to species
colnames(tax)[1] <- "species"
tax <- tax %>% as.matrix()

# reassign to phyloseq
x <- ps_FG
tax_table(x) <- tax_table(tax)

x <- aggregate_rare(x, level = "species", detection = 0, prevalence = 0)
x               
# 8 taxa and 140 samples


# remove "Unknown"
#allTaxa = taxa_names(x)
#badTaxa = c("Unknown")
#myTaxa <- allTaxa[!(allTaxa %in% badTaxa)]
#x <- prune_taxa(myTaxa, x)
#x
# 7 taxa and 140 samples

x_RA <- transform(x, 'compositional')

```

```{r}
#create data table
FG_df <-  psmelt(x_RA)
unique(FG_df$species)
```


```{r}

FG_df$species <- as.factor(FG_df$species)

for (i in levels(FG_df$species)) {
  df <- filter(FG_df, species == i)

  print(i)
  result = leveneTest(Abundance ~ sample_type, df)
  print(result)
  k <- kruskal.test(Abundance ~ sample_type, data = df)
  print(k)
  w <- pairwise.wilcox.test(df$Abundance, df$sample_type,
                     p.adjust.method = "BH")
  print(w)
  res.aov <- aov(Abundance ~ sample_type, data = df)
  aov <- summary(res.aov)
  print(aov)
  tukey <- TukeyHSD(res.aov)
  print(tukey)
  }

x <- FG_df %>% group_by(species, sample_type) %>%
  summarise(mean = mean(Abundance, na.rm = TRUE), se = (sd(Abundance, na.rm = TRUE)/sqrt(length((Abundance)))))
x
```


# 4. Trophic modes in forest deepest layer

```{r}
ps_x <- subset_samples(ps_FG, sample_type=="forest" & depth=="40...")
ps_x

ps_FG_Tm <- aggregate_rare(ps_x, level = "trophicMode", detection = 0, prevalence = 0)
ps_FG_Tm
# 8 taxa and 15 samples

# remove "Unknown"
#allTaxa = taxa_names(ps_FG_Tm)
#badTaxa = c("Unknown")
#myTaxa <- allTaxa[!(allTaxa %in% badTaxa)]
#ps_FG_Tm <- prune_taxa(myTaxa, ps_FG_Tm)
#ps_FG_Tm
# 7 taxa and 15 samples

ps_FG_Tm_RA <- microbiome::transform(ps_FG_Tm, "compositional")

FG_df <- psmelt(ps_FG_Tm_RA)

x <- FG_df %>%
  group_by(OTU) %>%
  summarise(mean = mean(Abundance, na.rm = TRUE), se = (sd(Abundance, na.rm = TRUE)/sqrt(length((Abundance)))))

x
```



# 5. Test symbiotrophic related trophic modes in soillayers

```{r}
# remove species

z <- as.data.frame(tax_table(ps_FG))

tax <- z[, 1:9]

tax <- tax %>%  
   mutate(FG = case_when(trophicMode == "Symbiotroph" | trophicMode == "Saprotroph-Symbiotroph" ~ "Symbiotroph-or-Saprotroph-Symbiotroph", trophicMode != "Symbiotroph" | trophicMode != "Saprotroph-Symbiotroph" ~ "Other"))

# rename FG to species
colnames(tax)[10] <- "species"

tax <- tax %>% as.matrix()

# reassign to phyloseq
x <- ps_FG
tax_table(x) <- tax_table(tax)

x <- aggregate_rare(x, level = "species", detection = 0, prevalence = 0)
x               
# 3 taxa and 140 samples


# remove "Unknown"
#allTaxa = taxa_names(x)
#badTaxa = c("Unknown")
#myTaxa <- allTaxa[!(allTaxa %in% badTaxa)]
#x <- prune_taxa(myTaxa, x)
#x
# 2 taxa and 140 samples

x_RA <- transform(x, 'compositional')

```


```{r}
#create data table
FG_df <-  psmelt(x_RA)
unique(FG_df$species)
```


```{r}


df <- filter(FG_df, species == "Symbiotroph-or-Saprotroph-Symbiotroph")
df <- filter(df, sample_type!="forest")

result = leveneTest(Abundance ~ depth, df)
print(result)

k <- kruskal.test(Abundance ~ depth, data = df)
print(k)
w <- pairwise.wilcox.test(df$Abundance, df$depth,
                     p.adjust.method = "BH")
print(w)



x <- df %>% group_by(depth) %>%
  summarise(mean = mean(Abundance, na.rm = TRUE), se = (sd(Abundance, na.rm = TRUE)/sqrt(length((Abundance)))))
x
```

# 6. Test symbiotrph related in 30-40 bewteen meadow, organic and conventional managements

```{r}
df <- filter(FG_df, species == "Symbiotroph-or-Saprotroph-Symbiotroph")
df <- filter(df, sample_type!="forest")
df <- filter(df, depth=="30...40")

result = leveneTest(Abundance ~ sample_type, df)
print(result)

res.aov <- aov(Abundance ~ sample_type, data = df)
aov <- summary(res.aov)
print(aov)
tukey <- TukeyHSD(res.aov)
print(tukey)

k <- kruskal.test(Abundance ~ sample_type, data = df)
print(k)
w <- pairwise.wilcox.test(df$Abundance, df$sample_type,
                     p.adjust.method = "BH")
print(w)


x <- df %>% group_by(sample_type) %>%
  summarise(mean = mean(Abundance, na.rm = TRUE), se = (sd(Abundance, na.rm = TRUE)/sqrt(length((Abundance)))))
x
```


# 7. Test symbiotrophs between soillayers

```{r}
# remove species

z <- as.data.frame(tax_table(ps_FG))

tax <- z[, 1:9]

# rename trophic mode to species
colnames(tax)[1] <- "species"

tax <- tax %>% as.matrix()

# reassign to phyloseq
x <- ps_FG
tax_table(x) <- tax_table(tax)

x <- aggregate_rare(x, level = "species", detection = 0, prevalence = 0)
x               
# 8 taxa and 140 samples


# remove "Unknown"
#allTaxa = taxa_names(x)
#badTaxa = c("Unknown")
#myTaxa <- allTaxa[!(allTaxa %in% badTaxa)]
#x <- prune_taxa(myTaxa, x)
#x
# 2 taxa and 140 samples

x_RA <- transform(x, 'compositional')

```


```{r}
#create data table
FG_df <-  psmelt(x_RA)
# remove forest
FG_df <- subset(FG_df, sample_type!="forest" | sample_type!="meadow")

unique(FG_df$species)
```


```{r}
df <- filter(FG_df, species == "Symbiotroph")

result = leveneTest(Abundance ~ depth, df)
print(result)

k <- kruskal.test(Abundance ~ depth, data = df)
print(k)
w <- pairwise.wilcox.test(df$Abundance, df$depth,
                     p.adjust.method = "BH")
print(w)



x <- df %>% group_by(depth) %>%
  summarise(mean = mean(Abundance, na.rm = TRUE), se = (sd(Abundance, na.rm = TRUE)/sqrt(length((Abundance)))))
x
```


# 8. Test saprotrophs between soillayers

```{r}
# remove species

z <- as.data.frame(tax_table(ps_FG))

tax <- z[, 1:9]

# rename trophic mode to species
colnames(tax)[1] <- "species"

tax <- tax %>% as.matrix()

# reassign to phyloseq
x <- ps_FG
tax_table(x) <- tax_table(tax)

x <- aggregate_rare(x, level = "species", detection = 0, prevalence = 0)
x               
# 8 taxa and 140 samples


# remove "Unknown"
#allTaxa = taxa_names(x)
#badTaxa = c("Unknown")
#myTaxa <- allTaxa[!(allTaxa %in% badTaxa)]
#x <- prune_taxa(myTaxa, x)
#x
# 2 taxa and 140 samples

x_RA <- transform(x, 'compositional')

```


```{r}
#create data table
FG_df <-  psmelt(x_RA)
# remove forest
FG_df <- subset(FG_df, sample_type!="forest" & sample_type!="meadow")

unique(FG_df$species)
```


```{r}
df <- filter(FG_df, species == "Saprotroph")

result = leveneTest(Abundance ~ depth, df)
print(result)

k <- kruskal.test(Abundance ~ depth, data = df)
print(k)
w <- pairwise.wilcox.test(df$Abundance, df$depth,
                     p.adjust.method = "BH")
print(w)



x <- df %>% group_by(depth) %>%
  summarise(mean = mean(Abundance, na.rm = TRUE), se = (sd(Abundance, na.rm = TRUE)/sqrt(length((Abundance)))))
x
```



# 9 Test Trophic modes in soillayers

```{r}
# remove species

z <- as.data.frame(tax_table(ps_FG))

tax <- z[, 1:9]


# rename trophic mode to species
colnames(tax)[1] <- "species"

tax <- tax %>% as.matrix()

# reassign to phyloseq
x <- ps_FG
tax_table(x) <- tax_table(tax)

x <- aggregate_rare(x, level = "species", detection = 0, prevalence = 0)
x               
# 8 taxa and 140 samples


# remove "Unknown"
#allTaxa = taxa_names(x)
#badTaxa = c("Unknown")
#myTaxa <- allTaxa[!(allTaxa %in% badTaxa)]
#x <- prune_taxa(myTaxa, x)
#x
# 2 taxa and 140 samples

x_RA <- transform(x, 'compositional')

```

```{r}
#create data table
FG_df <-  psmelt(x_RA)

unique(FG_df$species)
```

```{r}
FG_df <- filter(FG_df, sample_type != "forest")
FG_df <- droplevels(FG_df)

# make factor
FG_df$species <- as.factor(FG_df$species)
FG_df$depth <- as.factor(FG_df$depth)

#create data frame with 0 rows and 6 columns
df_test <- data.frame(matrix(ncol = 6))

#provide column names
colnames(df_test) <- c('sample_type', 'species', 'depth', 'LevenesP', 'ANOVAsOrKrusalsP', 'mean_se_cld')

for (i in levels(FG_df$depth)) {
  for (j in levels(FG_df$species)) {
    df <- filter(FG_df, depth == i & species == j)
    levene <- leveneTest(Abundance ~ sample_type, df)
    if (levene[1,3]>0.05){
      res.aov <- aov(Abundance ~ sample_type, data = df)
    res.aov2 <- summary(res.aov)
    res.aov2 <- res.aov2[[1]]
    df.res.aov <- data.frame(matrix(ncol = 3, nrow =3))
    colnames(df.res.aov) <- c('sample_type', 'LevenesP', 'ANOVAsOrKrusalsP')
    df.res.aov$sample_type <- levels(FG_df$sample_type)
    df.res.aov$LevenesP <- levene[1,3]
    df.res.aov$ANOVAsOrKrusalsP <- res.aov2[1,5]
    df.res.aov$LevenesP <- sprintf("%.3f", round(df.res.aov$LevenesP, digits = 3))
    df.res.aov$ANOVAsOrKrusalsP <- sprintf("%.3f", round(df.res.aov$ANOVAsOrKrusalsP, digits = 3))
    
    tukey <- glht(res.aov, linfct=mcp(sample_type="Tukey"))
    cld <- cld(tukey)
    cld <- cld[["mcletters"]][["Letters"]]
    cld <- as.data.frame(cld)
    cld$sample_type <- row.names(cld)
    row.names(cld) <- NULL
    } else {
      res.aov2 <- kruskal.test(Abundance ~ sample_type, data = df)
    df.res.aov <- data.frame(matrix(ncol = 3, nrow =3))
    colnames(df.res.aov) <- c('sample_type', 'LevenesP', 'ANOVAsOrKrusalsP')
    df.res.aov$sample_type <- levels(FG_df$sample_type)
    df.res.aov$LevenesP <- levene[1,3]
    df.res.aov$ANOVAsOrKrusalsP <- res.aov2[["p.value"]]
    df.res.aov$LevenesP <- sprintf("%.3f", round(df.res.aov$LevenesP, digits = 3))
    df.res.aov$ANOVAsOrKrusalsP <- sprintf("%.3f", round(df.res.aov$ANOVAsOrKrusalsP, digits = 3))
    wilcox.res <- pairwise.wilcox.test(df$Abundance, df$sample_type,
                     p.adjust.method = "BH")
    wilcox.res <- wilcox.res[["p.value"]]
    
    cld = fullPTable(wilcox.res)
    cld[is.na(cld)] <- 0
    cld <- multcompLetters(cld)
    cld <- cld[["Letters"]]
    cld <- as.data.frame(cld)
    cld$sample_type <- row.names(cld)
    row.names(cld) <- NULL
    }
    
    x <- df %>%
      dplyr::group_by(sample_type) %>%
      dplyr::summarise(mean = mean(Abundance, na.rm = TRUE), se = (sd(Abundance, na.rm = TRUE)/sqrt(length((Abundance)))))
    
    x$mean <- sprintf("%.3f", round(x$mean, digits = 3))
    x$se <- sprintf("%.3f", round(x$se, digits = 3))
    x <- cbind(species = 0, x)
    x$species <- j
    x$depth <- i
    
    output <- merge(x, df.res.aov, by='sample_type')
    output <- merge(output, cld, by='sample_type')
    
    output$mean_se = paste(output$mean, output$se, sep="±")
    output$mean_se_cld = paste(output$mean_se, output$cld, sep="")
    output <- output[, -c(3,4,8,9)]
    
    df_test <- rbind(df_test, output)
  }
}

```


```{r}
df_test <- na.omit(df_test)

setwd('C:\\Users\\lehakkin\\PhD\\Fungal_ITS_soil_vertical_profile')

#write.csv2(df_test, file = "test.csv")
```


