---
title: "Untitled"
author: "Laura_Hakkinen"
date: "2024-06-05"
output: html_document
---

# RESULTS STEP 10: Richness figures and statistical testing 

```{r}

library("ggpubr")
library('phyloseq')
library("cowplot")
library("dplyr")
library("ggplot2")
library("vegan")
library("microbiome")
library("tibble")
library(stringr)
library(reshape2)
library(tidyr)


setwd('C:\\Users\\lehakkin\\PhD\\Fungal_ITS_soil_vertical_profile')

load(file = 'ps_FG_with_NAs')#ps_FG
ps_FG

load('ps_FINAL')
ps

meta <- meta(ps)

```

# 1. Calculate FUNGuild richnesses


## 1.1. Saprotroph 

```{r}
x_sub <- subset_taxa(ps_FG, trophicMode %in% c("Saprotroph"))
x_sub
# 4842 taxa and 140 samples

# Save output of `plot_richness` as a variable
richness <- plot_richness(x_sub, x = "sample_type", measures = c("Observed"), color = "depth")

# get the data into data frame
richness_df <- richness$data

# the observed richness is in the column "value", lets change this to saprotroph_richness
colnames(richness_df)[30] <- "saprotroph_richness"
# and lets remove the "variable" and "se" columns
richness_df <- subset(richness_df, select = -variable)
richness_df <- subset(richness_df, select = -se)

FG_richness <- richness_df

```


## 1.2. Symbiotroph 

```{r}
x_sub <- subset_taxa(ps_FG, trophicMode %in% c("Symbiotroph"))
x_sub
# 784 taxa and 140 samples



# Save output of `plot_richness` as a variable
richness <- plot_richness(x_sub, x = "sample_type", measures = c("Observed"), color = "depth")

# get the data
richness_df <- richness$data

# the observed richness is in the column "value", lets change this to symbiotroph_richness
colnames(richness_df)[30] <- "symbiotroph_richness"
# and lets remove extra columns
richness_df <- richness_df[, c(1, 30)]

# combine
FG_richness <- dplyr::left_join(FG_richness, richness_df, by = "sampleID")
FG_richness <- subset(FG_richness, select = -samples)

# sampleID into rownames
rownames(FG_richness) <- FG_richness[,1]

```


## 1.3. Pathotroph  

```{r}
x_sub <- subset_taxa(ps_FG, trophicMode %in% c("Pathotroph"))
x_sub
# 1500 taxa and 140 samples



# Save output of `plot_richness` as a variable
richness <- plot_richness(x_sub, x = "sample_type", measures = c("Observed"), color = "depth")

# get the data
richness_df <- richness$data

# the observed richness is in the column "value", lets change this to pathotroph_richness
colnames(richness_df)[30] <- "pathotroph_richness"
# and lets remove extra columns
richness_df <- richness_df[, c(1, 30)]

# combine
FG_richness <- dplyr::left_join(FG_richness, richness_df, by = "sampleID")


# sampleID into rownames
rownames(FG_richness) <- FG_richness[,1]

```


# 2. Calculate AMF (Glomeromycota) Richness

Note! AMF richness calculated from FUNGuild is exactly the same (not shown here)!

```{r}
x_sub <- subset_taxa(ps, phylum %in% c("Glomeromycota"))
x_sub
# 263 taxa and 140 samples



# Save output of `plot_richness` as a variable
richness <- plot_richness(x_sub, x = "sample_type", measures = c("Observed"), color = "depth")

# get the data
richness_df <- richness$data

# the observed richness is in the column "value", lets change this to AMF_richness
colnames(richness_df)[31] <- "AMF_richness"
# and lets remove extra columns
richness_df <- richness_df[, c(1, 31)]

# combine
FG_richness <- dplyr::left_join(FG_richness, richness_df, by = "sampleID")

# sampleID into rownames
rownames(FG_richness) <- FG_richness[,1]

# I actually want the meta data from the ps rather than the ps_FG, so leyt's remove extra columns

FG_richness <- FG_richness[, c(1, 28:31)]

meta <- dplyr::left_join(meta, FG_richness, by = "sampleID")

# sampleID into rownames
rownames(meta) <- meta[,1]

```


# 3. Save new richness measures in meta data

```{r}
sample_data(ps) <- sample_data(meta)

setwd('C:\\Users\\lehakkin\\PhD\\Fungal_ITS_soil_vertical_profile')

save(ps, file = 'ps_FINAL')
ps
```


# 4. Richness plots

Note! The CLD letters that will be added to the fiugures are based on statistical test results done at step 6.

## 4.1. Fungal richness

```{r}

# create your own color palette for sample types
MyPalette <- c(forest = "#1167b1", meadow = "#fbc02d", organic = "#8a8a8a", conventional =  "#b71c1c")

# OTU richness
OTU_rich <- meta %>%
  group_by(sample_type, depth_numerical) %>%
  summarise(mean = mean(observed, na.rm = TRUE), se = (sd(observed, na.rm = TRUE)/sqrt(length((observed))))) %>%
  ungroup() %>%
  ggplot(aes(y=mean, depth_numerical, color=sample_type)) +
  geom_line(linetype = "dashed") +
  geom_point(size=3, position=position_dodge(1.2)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.5, position=position_dodge(1.2)) + coord_flip() + scale_x_reverse() + theme(plot.title = element_text(size = 20, hjust = 0.5)) + theme(panel.border = element_rect(colour = "black", fill=NA, size=0.5)) + theme_cowplot() +
  theme(axis.text = element_text(size=16),
        axis.title = element_text(size=16),
        legend.text = element_text(size=16),
        legend.title = element_text(size=16),
        title = element_text(size=18)) +
  scale_y_continuous(name="Fungal richness") +
  labs(x = "depth") + theme(legend.title = element_blank()) + scale_colour_manual(values = MyPalette)
print(OTU_rich)

```



Change to thousands of reads (x1000)

```{r}
rich_k <- meta %>% dplyr::mutate(richness_k = observed/1000 ) %>%
  group_by(sample_type, depth_numerical) %>%
  summarise(mean = mean(richness_k, na.rm = TRUE), se = (sd(richness_k, na.rm = TRUE)/sqrt(length((richness_k))))) %>%
  ungroup() %>%
  ggplot(aes(y=mean, depth_numerical, color=sample_type)) +
  geom_line(linetype = "dashed") +
  geom_point(size=3, position=position_dodge(1.2)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.5, position=position_dodge(1.2)) + coord_flip() + scale_x_reverse() + theme(plot.title = element_text(size = 20, hjust = 0.5)) + theme(panel.border = element_rect(colour = "black", fill=NA, size=0.5)) + theme_cowplot() +
  theme(axis.text = element_text(size=16),
        axis.title = element_text(size=16),
        legend.text = element_text(size=16),
        legend.title = element_text(size=16),
        title = element_text(size=18)) +
  scale_y_continuous(name="Fungal richness \n (x1000)") +
  labs(x = "depth") + theme(legend.title = element_blank()) + scale_colour_manual(values = MyPalette)

rich_k
```

### 4.1.1 change labels

```{r}
rich_k <- meta %>% dplyr::mutate(richness_k = observed/1000 ) %>%
  group_by(sample_type, depth_numerical) %>%
  summarise(mean = mean(richness_k, na.rm = TRUE), se = (sd(richness_k, na.rm = TRUE)/sqrt(length((richness_k))))) %>%
  ungroup() %>%
  ggplot(aes(y=mean, depth_numerical, color=sample_type)) +
  geom_line(linetype = "dashed") +
  geom_point(size=3, position=position_dodge(1.2)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.5, position=position_dodge(1.2)) + theme(plot.title = element_text(size = 20, hjust = 0.5)) + theme(panel.border = element_rect(colour = "black", fill=NA, size=0.5)) + theme_cowplot() +
  theme(axis.text = element_text(size=16),
        axis.title = element_text(size=16),
        legend.text = element_text(size=16),
        legend.title = element_text(size=16),
        title = element_text(size=18)) +
  scale_y_continuous(name="Fungal richness \n (x1000)") +
  labs(x = NULL) + theme(legend.title = element_blank()) + scale_colour_manual(values = MyPalette)

rich_k <- rich_k + scale_x_reverse(breaks = c(60, 35, 25, 15, 5), labels = c("40-80 cm", "30-40 cm", "20-30 cm", "10-20 cm", "0-10 cm"))

rich_k <- rich_k + coord_flip()
rich_k

```


```{r}
rich_k <- meta %>% dplyr::mutate(richness_k = observed/1000 ) %>%
  group_by(sample_type, depth_numerical) %>%
  summarise(mean = mean(richness_k, na.rm = TRUE), se = (sd(richness_k, na.rm = TRUE)/sqrt(length((richness_k))))) %>%
  ungroup() %>%
  ggplot(aes(y=mean, depth_numerical, color=sample_type)) +
  geom_line(linetype = "dashed") +
  geom_point(size=3, position=position_dodge(1.2)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.5, position=position_dodge(1.2)) + theme(plot.title = element_text(size = 20, hjust = 0.5)) + theme(panel.border = element_rect(colour = "black", fill=NA, size=0.5)) + theme_cowplot() +
  theme(axis.text = element_text(size=16),
        axis.title = element_text(size=16),
        legend.text = element_text(size=16),
        legend.title = element_text(size=16),
        title = element_text(size=18)) +
  scale_y_continuous(name="Fungal richness \n (x1000)") +
  labs(x = NULL) + theme(legend.title = element_blank()) + scale_colour_manual(values = MyPalette)

rich_k <- rich_k + scale_x_reverse(breaks = c(60, 35, 25, 15, 5), labels = c("40-80 cm", "30-40 cm", "20-30 cm", "10-20 cm", "0-10 cm"))

rich_k <- rich_k +
  annotate("text", x=1.5, y=1.5, label= "b", col="#1167b1", size=3) +
  annotate("text", x=1.5, y=1.1, label= "a", col="#fbc02d", size=3) +
  annotate("text", x=1.5, y=3, label= "c", col="#8a8a8a", size=3) +
  annotate("text", x=1.5, y=2.6, label= "c", col="#b71c1c", size=3) +
  annotate("text", x=11, y=1.4, label= "a", col="#1167b1", size=3) +
  annotate("text", x=11, y=2.0, label= "ab", col="#fbc02d", size=3) +
  annotate("text", x=11, y=2.55, label= "ab", col="#8a8a8a", size=3) +
  annotate("text", x=11, y=3.0, label= "b", col="#b71c1c", size=3) +
  annotate("text", x=25, y=0.2, label= "(ns)", size=3) +
  annotate("text", x=35, y=1.2, label= "(ns)", size=3) +
  annotate("text", x=59, y=0.7, label= "ab", col="#1167b1", size=3) +
  annotate("text", x=59, y=1.2, label= "ab", col="#fbc02d", size=3) +
  annotate("text", x=59, y=1.45, label= "b", col="#8a8a8a", size=3) +
  annotate("text", x=59, y=0.95, label= "a", col="#b71c1c", size=3)

rich_k <- rich_k + coord_flip()
rich_k
```



## 4.2. AMF richness

```{r}
gm_rich <- meta %>%
  group_by(sample_type, depth_numerical) %>%
  summarise(mean = mean(AMF_richness, na.rm = TRUE), se = (sd(AMF_richness, na.rm = TRUE)/sqrt(length((AMF_richness))))) %>%
  ungroup() %>%
  ggplot(aes(y=mean, depth_numerical, color=sample_type)) +
  geom_line(linetype = "dashed") +
  geom_point(size=3, position=position_dodge(1.2)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.5, position=position_dodge(1.2)) + coord_flip() + scale_x_reverse() + theme(plot.title = element_text(size = 20, hjust = 0.5)) + theme(panel.border = element_rect(colour = "black", fill=NA, size=0.5)) + theme_cowplot() +
  theme(axis.text = element_text(size=16),
        axis.title = element_text(size=16),
        legend.text = element_text(size=16),
        legend.title = element_text(size=16),
        title = element_text(size=18)) +
  scale_y_continuous(name="AMF richness \n") +
  labs(x = "depth") + theme(legend.title = element_blank()) + scale_colour_manual(values = MyPalette)

gm_rich
```

### 4.2.1 change labels

```{r}
gm_rich <- meta %>%
  group_by(sample_type, depth_numerical) %>%
  summarise(mean = mean(AMF_richness, na.rm = TRUE), se = (sd(AMF_richness, na.rm = TRUE)/sqrt(length((AMF_richness))))) %>%
  ungroup() %>%
  ggplot(aes(y=mean, depth_numerical, color=sample_type)) +
  geom_line(linetype = "dashed") +
  geom_point(size=3, position=position_dodge(1.2)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.5, position=position_dodge(1.2)) + theme(plot.title = element_text(size = 20, hjust = 0.5)) + theme(panel.border = element_rect(colour = "black", fill=NA, size=0.5)) + theme_cowplot() +
  theme(axis.text = element_text(size=16),
        axis.title = element_text(size=16),
        legend.text = element_text(size=16),
        legend.title = element_text(size=16),
        title = element_text(size=18)) +
  scale_y_continuous(name="AMF richness \n") +
  labs(x = NULL) + theme(legend.title = element_blank()) + scale_colour_manual(values = MyPalette)

gm_rich <- gm_rich + scale_x_reverse(breaks = c(60, 35, 25, 15, 5), labels = c("40-80 cm", "30-40 cm", "20-30 cm", "10-20 cm", "0-10 cm"))


gm_rich <- gm_rich +
  annotate("text", x=1.5, y=7, label= "a", col="#1167b1", size=3) +
  annotate("text", x=1.5, y=12, label= "ab", col="#fbc02d", size=3) +
  annotate("text", x=1.5, y=17, label= "ab", col="#8a8a8a", size=3) +
  annotate("text", x=1.5, y=22, label= "b", col="#b71c1c", size=3) +
  annotate("text", x=11, y=11, label= "a", col="#1167b1", size=3) +
  annotate("text", x=11, y=30, label= "b", col="#fbc02d", size=3) +
  annotate("text", x=11, y=23, label= "ab", col="#8a8a8a", size=3) +
  annotate("text", x=11, y=27, label= "b", col="#b71c1c", size=3) +
  annotate("text", x=21, y=9, label= "a", col="#1167b1", size=3) +
  annotate("text", x=21, y=37, label= "b", col="#fbc02d", size=3) +
  annotate("text", x=21, y=27, label= "ab", col="#8a8a8a", size=3) +
  annotate("text", x=21, y=20, label= "a", col="#b71c1c", size=3) +
  annotate("text", x=31, y=8, label= "ab", col="#1167b1", size=3) +
  annotate("text", x=31, y=34, label= "b", col="#fbc02d", size=3) +
  annotate("text", x=31, y=17, label= "ab", col="#8a8a8a", size=3) +
  annotate("text", x=31, y=4, label= "a", col="#b71c1c", size=3) +
  annotate("text", x=59, y=13, label= "(ns)", size=3)

gm_rich <- gm_rich + coord_flip()
gm_rich
```


## 4.3. Saprotroph richness

```{r}
Saprotroph_rich <- meta %>%
  group_by(sample_type, depth_numerical) %>%
  summarise(mean = mean(saprotroph_richness, na.rm = TRUE), se = (sd(saprotroph_richness, na.rm = TRUE)/sqrt(length((saprotroph_richness))))) %>%
  ungroup() %>%
  ggplot(aes(y=mean, depth_numerical, color=sample_type)) +
  geom_line(linetype = "dashed") +
  geom_point(size=3, position=position_dodge(1.2)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.5, position=position_dodge(1.2)) + coord_flip() + scale_x_reverse() + theme(plot.title = element_text(size = 20, hjust = 0.5)) + theme(panel.border = element_rect(colour = "black", fill=NA, size=0.5)) + theme_cowplot() +
  theme(axis.text = element_text(size=16),
        axis.title = element_text(size=16),
        legend.text = element_text(size=16),
        legend.title = element_text(size=16),
        title = element_text(size=18)) +
  scale_y_continuous(name="Saprotroph richness \n") +
  labs(x = "depth") + theme(legend.title = element_blank()) + scale_colour_manual(values = MyPalette)


Saprotroph_rich

```

### 4.3.1 change labels

```{r}
Saprotroph_rich <- meta %>%
  group_by(sample_type, depth_numerical) %>%
  summarise(mean = mean(saprotroph_richness, na.rm = TRUE), se = (sd(saprotroph_richness, na.rm = TRUE)/sqrt(length((saprotroph_richness))))) %>%
  ungroup() %>%
  ggplot(aes(y=mean, depth_numerical, color=sample_type)) +
  geom_line(linetype = "dashed") +
  geom_point(size=3, position=position_dodge(1.2)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.5, position=position_dodge(1.2)) + theme(plot.title = element_text(size = 20, hjust = 0.5)) + theme(panel.border = element_rect(colour = "black", fill=NA, size=0.5)) + theme_cowplot() +
  theme(axis.text = element_text(size=16),
        axis.title = element_text(size=16),
        legend.text = element_text(size=16),
        legend.title = element_text(size=16),
        title = element_text(size=18)) +
  scale_y_continuous(name="Saprotroph richness \n") +
  labs(x = NULL) + theme(legend.title = element_blank()) + scale_colour_manual(values = MyPalette)

Saprotroph_rich <- Saprotroph_rich + scale_x_reverse(breaks = c(60, 35, 25, 15, 5), labels = c("40-80 cm", "30-40 cm", "20-30 cm", "10-20 cm", "0-10 cm"))

Saprotroph_rich <- Saprotroph_rich +
  annotate("text", x=1.5, y=450, label= "a", col="#1167b1", size=3) +
  annotate("text", x=1.5, y=550, label= "a", col="#fbc02d", size=3) +
  annotate("text", x=1.5, y=830, label= "b", col="#8a8a8a", size=3) +
  annotate("text", x=1.5, y=690, label= "ab", col="#b71c1c", size=3) +
  annotate("text", x=11, y=490, label= "a", col="#1167b1", size=3) +
  annotate("text", x=11, y=900, label= "b", col="#fbc02d", size=3) +
  annotate("text", x=11, y=680, label= "ab", col="#8a8a8a", size=3) +
  annotate("text", x=11, y=800, label= "ab", col="#b71c1c", size=3) +
  annotate("text", x=21, y=220, label= "a", col="#1167b1", size=3) +
  annotate("text", x=21, y=770, label= "b", col="#fbc02d", size=3) +
  annotate("text", x=21, y=600, label= "ab", col="#8a8a8a", size=3) +
  annotate("text", x=21, y=450, label= "ab", col="#b71c1c", size=3) +
  annotate("text", x=35, y=450, label= "(ns)", size=3) +
  annotate("text", x=59, y=200, label= "(ns)", size=3)

Saprotroph_rich <- Saprotroph_rich + coord_flip()
Saprotroph_rich

```



## 4.4. Symbiotroph richness

```{r}
symb_rich <- meta %>%
  group_by(sample_type, depth_numerical) %>%
  summarise(mean = mean(symbiotroph_richness, na.rm = TRUE), se = (sd(symbiotroph_richness, na.rm = TRUE)/sqrt(length((symbiotroph_richness))))) %>%
  ungroup() %>%
  ggplot(aes(y=mean, depth_numerical, color=sample_type)) +
  geom_line(linetype = "dashed") +
  geom_point(size=3, position=position_dodge(1.2)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.5, position=position_dodge(1.2)) + coord_flip() + scale_x_reverse() + theme(plot.title = element_text(size = 20, hjust = 0.5)) + theme(panel.border = element_rect(colour = "black", fill=NA, size=0.5)) + theme_cowplot() +
  theme(axis.text = element_text(size=16),
        axis.title = element_text(size=16),
        legend.text = element_text(size=16),
        legend.title = element_text(size=16),
        title = element_text(size=18)) +
  scale_y_continuous(name="Symbiotroph richness \n") +
  labs(x = "depth") + theme(legend.title = element_blank()) + scale_colour_manual(values = MyPalette)


symb_rich
```

### 4.4.1 change labels

```{r}
symb_rich <- meta %>%
  group_by(sample_type, depth_numerical) %>%
  summarise(mean = mean(symbiotroph_richness, na.rm = TRUE), se = (sd(symbiotroph_richness, na.rm = TRUE)/sqrt(length((symbiotroph_richness))))) %>%
  ungroup() %>%
  ggplot(aes(y=mean, depth_numerical, color=sample_type)) +
  geom_line(linetype = "dashed") +
  geom_point(size=3, position=position_dodge(1.2)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.5, position=position_dodge(1.2)) + theme(plot.title = element_text(size = 20, hjust = 0.5)) + theme(panel.border = element_rect(colour = "black", fill=NA, size=0.5)) + theme_cowplot() +
  theme(axis.text = element_text(size=16),
        axis.title = element_text(size=16),
        legend.text = element_text(size=16),
        legend.title = element_text(size=16),
        title = element_text(size=18)) +
  scale_y_continuous(name="Symbiotroph richness \n") +
  labs(x = NULL) + theme(legend.title = element_blank()) + scale_colour_manual(values = MyPalette)

symb_rich <- symb_rich + scale_x_reverse(breaks = c(60, 35, 25, 15, 5), labels = c("40-80 cm", "30-40 cm", "20-30 cm", "10-20 cm", "0-10 cm"))

symb_rich <- symb_rich + coord_flip()
symb_rich

```


## 4.5. Pathotroph richness

```{r}
path_rich <- meta %>%
  group_by(sample_type, depth_numerical) %>%
  summarise(mean = mean(pathotroph_richness, na.rm = TRUE), se = (sd(pathotroph_richness, na.rm = TRUE)/sqrt(length((pathotroph_richness))))) %>%
  ungroup() %>%
  ggplot(aes(y=mean, depth_numerical, color=sample_type)) +
  geom_line(linetype = "dashed") +
  geom_point(size=3, position=position_dodge(1.2)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.5, position=position_dodge(1.2)) + coord_flip() + scale_x_reverse() + theme(plot.title = element_text(size = 20, hjust = 0.5)) + theme(panel.border = element_rect(colour = "black", fill=NA, size=0.5)) + theme_cowplot() +
  theme(axis.text = element_text(size=16),
        axis.title = element_text(size=16),
        legend.text = element_text(size=16),
        legend.title = element_text(size=16),
        title = element_text(size=18)) +
  scale_y_continuous(name="Pathotroph richness \n") +
  labs(x = "depth") + theme(legend.title = element_blank()) + scale_colour_manual(values = MyPalette)


path_rich
```

### 4.5.1 change labels

```{r}
path_rich <- meta %>%
  group_by(sample_type, depth_numerical) %>%
  summarise(mean = mean(pathotroph_richness, na.rm = TRUE), se = (sd(pathotroph_richness, na.rm = TRUE)/sqrt(length((pathotroph_richness))))) %>%
  ungroup() %>%
  ggplot(aes(y=mean, depth_numerical, color=sample_type)) +
  geom_line(linetype = "dashed") +
  geom_point(size=3, position=position_dodge(1.2)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.5, position=position_dodge(1.2)) + theme(plot.title = element_text(size = 20, hjust = 0.5)) + theme(panel.border = element_rect(colour = "black", fill=NA, size=0.5)) + theme_cowplot() +
  theme(axis.text = element_text(size=16),
        axis.title = element_text(size=16),
        legend.text = element_text(size=16),
        legend.title = element_text(size=16),
        title = element_text(size=18)) +
  scale_y_continuous(name="Pathotroph richness \n") +
  labs(x = NULL) + theme(legend.title = element_blank()) + scale_colour_manual(values = MyPalette)

path_rich <- path_rich + scale_x_reverse(breaks = c(60, 35, 25, 15, 5), labels = c("40-80 cm", "30-40 cm", "20-30 cm", "10-20 cm", "0-10 cm"))

path_rich <- path_rich +
  annotate("text", x=1.5, y=90, label= "a", col="#1167b1", size=3) +
  annotate("text", x=1.5, y=60, label= "a", col="#fbc02d", size=3) +
  annotate("text", x=1.5, y=310, label= "b", col="#8a8a8a", size=3) +
  annotate("text", x=1.5, y=330, label= "b", col="#b71c1c", size=3) +
  annotate("text", x=11, y=56, label= "a", col="#1167b1", size=3) +
  annotate("text", x=11, y=118, label= "a", col="#fbc02d", size=3) +
  annotate("text", x=11, y=275, label= "b", col="#8a8a8a", size=3) +
  annotate("text", x=11, y=331, label= "b", col="#b71c1c", size=3) +
  annotate("text", x=25, y=250, label= "(ns)", size=3) +
  annotate("text", x=35, y=70, label= "(ns)", size=3) +
  annotate("text", x=59, y=50, label= "(ns)", size=3)

path_rich <- path_rich + coord_flip()
path_rich

```



# 5. Combine selected figures

```{r fig.dim = c(14, 6)}

figure <- ggarrange(rich_k, gm_rich + rremove("ylab") + rremove("y.text") + rremove("y.axis") + rremove("y.ticks"), Saprotroph_rich + rremove("ylab") + rremove("y.text") + rremove("y.axis") + rremove("y.ticks"), path_rich + rremove("ylab") + rremove("y.text") + rremove("y.axis") + rremove("y.ticks"),
                    labels = c("A", "B", "C", "D"),
                    ncol = 4, nrow = 1, common.legend = TRUE, legend = "right", widths = c(1.25, 0.9, 0.9, 0.9))

figure
```


# 6. Test richnes between management types in all soil layers

```{r}
library("multcomp")
```


Change depth and richness measure accordingly

Test these:
- observed
- AMF_richness
- saprotroph_richness
- pathotroph_richness

```{r}

library(car)

means_and_ses <- list()

meta$depth <- as.factor(meta$depth)
rich <-c("observed", "AMF_richness", "saprotroph_richness", "pathotroph_richness")

for (i in rich) {
  for (j in levels(meta$depth)) {
    df <- filter(meta, depth==j)
    
     print(i)
     print(j)
     # Levene's test for homogeneity of variance
    result <- leveneTest(df[[i]] ~ sample_type, data = df)
    print(result)
     
    # Kruskal-Wallis test
    k <- kruskal.test(df[[i]] ~ sample_type, data = df)
    print(k)
    
    # Pairwise Wilcoxon test
    w <- pairwise.wilcox.test(df[[i]], df$sample_type, p.adjust.method = "BH")
    print(w)
    
     # ANOVA
    res.aov <- aov(df[[i]] ~ sample_type, data = df)
    aov_summary <- summary(res.aov)
    print(aov_summary)
    
    # Tukey's HSD test
    tukey <- TukeyHSD(res.aov)
    print(tukey)
    
    # Group by and summarize
    mean_and_se <- df %>%
      group_by(sample_type) %>%
      summarise(mean = mean(!!sym(i), na.rm = TRUE), se = sd(!!sym(i), na.rm = TRUE) / sqrt(n()))
  
    # Store the result in the list with a descriptive name
    result_name <- paste("depth", j, "diversity", i, sep = "_")
    means_and_ses[[result_name]] <- mean_and_se
    
  }
}

# Combine list into a single dataframe
combined_df <- bind_rows(lapply(names(means_and_ses), function(name) {
  df <- means_and_ses[[name]]
  df$result_name <- name
  return(df)
}), .id = "id")

# View the combined dataframe
print(combined_df)

```



```{r}
# save the mean and se values

setwd('C:\\Users\\lehakkin\\PhD\\Fungal_ITS_soil_vertical_profile')

#write.csv2(combined_df, file = "Richness_mean_and_ses.csv")
```






