---
title: "Untitled"
author: "Laura_Hakkinen"
date: "2024-06-26"
output: html_document
---

# RESULTS STEPS 5 and 6:HEATMAPS with genera and COMPOSITION PLOTS for phyla, class and FUNGuild

# A) HEATMAP for meadow, organic and conventional

## 1. load packages and data

```{r}

library('phyloseq')
library("cowplot")
library("dplyr")
library("ggplot2")
library("vegan")
library("microbiome")
library("tibble")


setwd('C:\\Users\\lehakkin\\PhD\\Fungal_ITS_soil_vertical_profile')

load('ps_FINAL')
ps


# Pick relative abundances (compositional) and sample metadata 
ps_RA <- microbiome::transform(ps, "compositional")
ps_RA_nf <- subset_samples(ps_RA, sample_type!="forest")

# You have only submitted samples, not OTUs. Define prevalence of each taxa
# (in how many samples did each taxa appear at least once)
prev0 = apply(X = otu_table(ps_RA_nf),
              MARGIN = ifelse(taxa_are_rows(ps_RA_nf), yes = 1, no = 2),
              FUN = function(x){sum(x > 0)})

# Execute prevalence filter, using `prune_taxa()` function
ps_RA_nf = prune_taxa((prev0 > 0), ps_RA_nf)
ps_RA_nf

meta_nf <- meta(ps_RA_nf)
```


# 2. Overview of the heatmap analysis

I will do heatmap for fungal genera using only meadow, organic and conventional soils (without forest).
I will use RA values
For the final heatmap figure the RA values will be standardizes (z-transformed)

1) First pick the 20 most abundant genera in each soil type depth
2) Do multiple testing between sample_type_depths that make sense (do not test between e.g. meadow deep and organic topsoil)
3) Make a HEATMAP of the significant taxa only

!! NOTE!! I will also include taxa that are not classified at genus level in the heatmap

# 3. at genus level

```{r}

ps_genus <- aggregate_rare(ps, level = 'genus', detection = 0/100, prevalence = 0/140)
ps_genus


ps_genus_RA <- microbiome::transform(ps_genus, "compositional")

ps_genus_RA_pruned <- subset_samples(ps_genus_RA, sample_type!="forest")
ps_genus_RA_pruned
```


# 4. Pick 20 most abund taxa in each soil management layer

```{r}
# sample wise filtering according to most abund. genera

# Initialize an empty list to store the taxa
abund.taxa <- list()

for (i in meta_nf$sample_type) {
  for (j in meta_nf$depth) {
    x <- sample_names(sample_data(ps_genus_RA_pruned)[sample_data(ps_genus_RA_pruned)$sample_type == i & sample_data(ps_genus_RA_pruned)$depth == j,])

# Calculate taxa mean of the selected samples
top20 <- head(sort(rowMeans(otu_table(ps_genus_RA_pruned)[,x]), decreasing = TRUE), 20)

result_name <- paste("sample_type", i, "depth", j, sep = "_")

abund.taxa[[result_name]] <- top20
  }
}

```



# 5. Combine picked taxa

```{r}
management_layer <- c(names(abund.taxa))

all_top20 <- c()

for (i in management_layer) {
  top20 <- c(names(abund.taxa[[i]]))
  all_top20 <- c(all_top20, top20)

}

all_top20_unique <- unique(all_top20)
length(all_top20_unique)

all_top20_unique_df <- as.data.frame(all_top20_unique)

```


# 6. Make a phyloseq of picked taxa

```{r}
# first aggregate to genus
ps_RA_nf_genus <- aggregate_rare(ps_RA_nf, level = "genus", detection = 0/140, prevalence = 0/100)

# filter by taxa name
ps_RA_nf_genus_pruned <- prune_taxa(all_top20_unique, ps_RA_nf_genus)
ps_RA_nf_genus_pruned

ps_genus_nf_HETAMAP <- ps_RA_nf_genus_pruned

```


# 7. Add sample_type_depth to meta


```{r}
meta_nf$sample_type_depth <- NA

meta_nf$sample_type_depth[meta_nf$depth_numerical==5 & meta_nf$sample_type=="meadow"]<-"0...10_meadow"
meta_nf$sample_type_depth[meta_nf$depth_numerical==5 & meta_nf$sample_type=="organic"]<-"0...10_organic"
meta_nf$sample_type_depth[meta_nf$depth_numerical==5 & meta_nf$sample_type=="conventional"]<-"0...10_conventional"

meta_nf$sample_type_depth[meta_nf$depth_numerical==15 & meta_nf$sample_type=="meadow"]<-"10...20_meadow"
meta_nf$sample_type_depth[meta_nf$depth_numerical==15 & meta_nf$sample_type=="organic"]<-"10...20_organic"
meta_nf$sample_type_depth[meta_nf$depth_numerical==15 & meta_nf$sample_type=="conventional"]<-"10...20_conventional"

meta_nf$sample_type_depth[meta_nf$depth_numerical==25 & meta_nf$sample_type=="meadow"]<-"20...30_meadow"
meta_nf$sample_type_depth[meta_nf$depth_numerical==25 & meta_nf$sample_type=="organic"]<-"20...30_organic"
meta_nf$sample_type_depth[meta_nf$depth_numerical==25 & meta_nf$sample_type=="conventional"]<-"20...30_conventional"

meta_nf$sample_type_depth[meta_nf$depth_numerical==35 & meta_nf$sample_type=="meadow"]<-"30...40_meadow"
meta_nf$sample_type_depth[meta_nf$depth_numerical==35 & meta_nf$sample_type=="organic"]<-"30...40_organic"
meta_nf$sample_type_depth[meta_nf$depth_numerical==35 & meta_nf$sample_type=="conventional"]<-"30...40_conventional"

meta_nf$sample_type_depth[meta_nf$depth_numerical==60 & meta_nf$sample_type=="meadow"]<-"40...80_meadow"
meta_nf$sample_type_depth[meta_nf$depth_numerical==60 & meta_nf$sample_type=="organic"]<-"40...80_organic"
meta_nf$sample_type_depth[meta_nf$depth_numerical==60 & meta_nf$sample_type=="conventional"]<-"40...80_conventional"


sample_type_depth <- unique(meta_nf$sample_type_depth)
sample_type_depth

# add the new meta_nf to ps_RA_nf_genus_pruned
sample_data(ps_RA_nf_genus_pruned) <- sample_data(meta_nf)

```


# 8. Test abundance differences

lets test separately for the ones that make sense, for example, no point testing between organic 0...10 and conventional 10...20, but rather 0...10 for both soil types and then within organic between 0...10 and 10...20

NOTE! At the end of chunk I  do p-value adjustment ("BH") for all comparisons in the chunk


## 8.1. 0...10 all soil types

```{r}
library("data.table")
library("rstatix")

subset <- subset_samples(ps_RA_nf_genus_pruned, sample_type_depth=="0...10_meadow" | sample_type_depth=="0...10_conventional" | sample_type_depth=="0...10_organic")
subset

#create data table
melt_df <-  psmelt(subset)
# Make sample_type_depth into a factor
melt_df$sample_type_depth <- factor(melt_df$sample_type_depth)

pval.list <- list()

for (i in all_top20_unique)
{
  # data.frame with the selected taxonomic group
  df <- filter(melt_df, genus==i)
  x <- pairwise.wilcox.test(df$Abundance, df$sample_type_depth,
                            p.adjust.method = NULL)
  x <- as.data.frame(x[["p.value"]])
  x$genus <- i
  x$comparison <- rownames(x)
  pval.list[[i]] <- x
}

pval.list_df1 <- rbindlist(pval.list)
pval.list_df1 <- as.data.frame(pval.list_df1)
# make into long format: where the new column  called "vs" contains the sample sample_type_depth comparison, and new column "p_value" contains the p_values
pval.list_df1 <- gather(pval.list_df1, vs, p_value, 1:2, factor_key=TRUE)
# adjust p values
pval.list_df1 <- adjust_pvalue(pval.list_df1, p.col = "p_value", output.col = "adj_p", method = "BH")

```


## 8.2. 10...20 all soil types

```{r}

subset <- subset_samples(ps_RA_nf_genus_pruned, sample_type_depth=="10...20_meadow" | sample_type_depth=="10...20_conventional" | sample_type_depth=="10...20_organic")
subset

#create data table
melt_df <-  psmelt(subset)

pval.list <- list()

for (i in all_top20_unique)
{
  # data.frame with the selected taxonomic group
  df <- filter(melt_df, genus==i)
  x <- pairwise.wilcox.test(df$Abundance, df$sample_type_depth,
                            p.adjust.method = "BH")
  x <- as.data.frame(x[["p.value"]])
  x$genus <- i
  x$comparison <- rownames(x)
  pval.list[[i]] <- x
}

pval.list_df2 <- rbindlist(pval.list)
pval.list_df2 <- as.data.frame(pval.list_df2)
# make into long format: where the new column  called "vs" contains the sample sample_type_depth comparison, and new column "p_value" contains the p_values
pval.list_df2 <- gather(pval.list_df2, vs, p_value, 1:2, factor_key=TRUE)
# adjust p values
pval.list_df2 <- adjust_pvalue(pval.list_df2, p.col = "p_value", output.col = "adj_p", method = "BH")
```


## 8.3. 20...30 all soil types

```{r}

subset <- subset_samples(ps_RA_nf_genus_pruned, sample_type_depth=="20...30_meadow" | sample_type_depth=="20...30_conventional" | sample_type_depth=="20...30_organic")
subset

#create data table
melt_df <-  psmelt(subset)

pval.list <- list()

for (i in all_top20_unique)
{
  # data.frame with the selected taxonomic group
  df <- filter(melt_df, genus==i)
  x <- pairwise.wilcox.test(df$Abundance, df$sample_type_depth,
                            p.adjust.method = "BH")
  x <- as.data.frame(x[["p.value"]])
  x$genus <- i
  x$comparison <- rownames(x)
  pval.list[[i]] <- x
}

pval.list_df3 <- rbindlist(pval.list)
pval.list_df3 <- as.data.frame(pval.list_df3)
# make into long format: where the new column  called "vs" contains the sample sample_type_depth comparison, and new column "p_value" contains the p_values
pval.list_df3 <- gather(pval.list_df3, vs, p_value, 1:2, factor_key=TRUE)
# adjust p values
pval.list_df3 <- adjust_pvalue(pval.list_df3, p.col = "p_value", output.col = "adj_p", method = "BH")
```


## 8.4. 30...40 all soil types

```{r}

subset <- subset_samples(ps_RA_nf_genus_pruned, sample_type_depth=="30...40_meadow" | sample_type_depth=="30...40_conventional" | sample_type_depth=="30...40_organic")
subset

#create data table
melt_df <-  psmelt(subset)

pval.list <- list()

for (i in all_top20_unique)
{
  # data.frame with the selected taxonomic group
  df <- filter(melt_df, genus==i)
  x <- pairwise.wilcox.test(df$Abundance, df$sample_type_depth,
                            p.adjust.method = "BH")
  x <- as.data.frame(x[["p.value"]])
  x$genus <- i
  x$comparison <- rownames(x)
  pval.list[[i]] <- x
}

pval.list_df4 <- rbindlist(pval.list)
pval.list_df4 <- as.data.frame(pval.list_df4)
# make into long format: where the new column  called "vs" contains the sample sample_type_depth comparison, and new column "p_value" contains the p_values
pval.list_df4 <- gather(pval.list_df4, vs, p_value, 1:2, factor_key=TRUE)
# adjust p values
pval.list_df4 <- adjust_pvalue(pval.list_df4, p.col = "p_value", output.col = "adj_p", method = "BH")
```


## 8.5. 40...80 all soil types

```{r}

subset <- subset_samples(ps_RA_nf_genus_pruned, sample_type_depth=="40...80_meadow" | sample_type_depth=="40...80_conventional" | sample_type_depth=="40...80_organic")
subset

#create data table
melt_df <-  psmelt(subset)

pval.list <- list()

for (i in all_top20_unique)
{
  # data.frame with the selected taxonomic group
  df <- filter(melt_df, genus==i)
  x <- pairwise.wilcox.test(df$Abundance, df$sample_type_depth,
                            p.adjust.method = "BH")
  x <- as.data.frame(x[["p.value"]])
  x$genus <- i
  x$comparison <- rownames(x)
  pval.list[[i]] <- x
}

pval.list_df5 <- rbindlist(pval.list)
pval.list_df5 <- as.data.frame(pval.list_df5)
# make into long format: where the new column  called "vs" contains the sample sample_type_depth comparison, and new column "p_value" contains the p_values
pval.list_df5 <- gather(pval.list_df5, vs, p_value, 1:2, factor_key=TRUE)
# adjust p values
pval.list_df5 <- adjust_pvalue(pval.list_df5, p.col = "p_value", output.col = "adj_p", method = "BH")
```


## 8.6. only organic 0-20

```{r}

subset <- subset_samples(ps_RA_nf_genus_pruned, sample_type_depth=="0...10_organic" | sample_type_depth=="10...20_organic")
subset

#create data table
melt_df <-  psmelt(subset)

pval.list <- list()

for (i in all_top20_unique)
{
  # data.frame with the selected taxonomic group
  df <- filter(melt_df, genus==i)
  x <- pairwise.wilcox.test(df$Abundance, df$sample_type_depth,
                            p.adjust.method = "BH")
  x <- as.data.frame(x[["p.value"]])
  x$genus <- i
  x$comparison <- rownames(x)
  pval.list[[i]] <- x
}

pval.list_df6 <- rbindlist(pval.list)
pval.list_df6 <- as.data.frame(pval.list_df6)
# make into long format: where the new column  called "vs" contains the sample sample_type_depth comparison, and new column "p_value" contains the p_values
pval.list_df6 <- gather(pval.list_df6, vs, p_value, 1, factor_key=TRUE)
# adjust p values
pval.list_df6 <- adjust_pvalue(pval.list_df6, p.col = "p_value", output.col = "adj_p", method = "BH")
```


## 8.7. only organic 10-30

```{r}

subset <- subset_samples(ps_RA_nf_genus_pruned, sample_type_depth=="10...20_organic" | sample_type_depth=="20...30_organic")
subset

#create data table
melt_df <-  psmelt(subset)

pval.list <- list()

for (i in all_top20_unique)
{
  # data.frame with the selected taxonomic group
  df <- filter(melt_df, genus==i)
  x <- pairwise.wilcox.test(df$Abundance, df$sample_type_depth,
                            p.adjust.method = "BH")
  x <- as.data.frame(x[["p.value"]])
  x$genus <- i
  x$comparison <- rownames(x)
  pval.list[[i]] <- x
}

pval.list_df7 <- rbindlist(pval.list)
pval.list_df7 <- as.data.frame(pval.list_df7)
# make into long format: where the new column  called "vs" contains the sample sample_type_depth comparison, and new column "p_value" contains the p_values
pval.list_df7 <- gather(pval.list_df7, vs, p_value, 1, factor_key=TRUE)
# adjust p values
pval.list_df7 <- adjust_pvalue(pval.list_df7, p.col = "p_value", output.col = "adj_p", method = "BH")
```

## 8.8. only organic 20-40

```{r}

subset <- subset_samples(ps_RA_nf_genus_pruned, sample_type_depth=="20...30_organic" | sample_type_depth=="30...40_organic")
subset

#create data table
melt_df <-  psmelt(subset)

pval.list <- list()

for (i in all_top20_unique)
{
  # data.frame with the selected taxonomic group
  df <- filter(melt_df, genus==i)
  x <- pairwise.wilcox.test(df$Abundance, df$sample_type_depth,
                            p.adjust.method = "BH")
  x <- as.data.frame(x[["p.value"]])
  x$genus <- i
  x$comparison <- rownames(x)
  pval.list[[i]] <- x
}

pval.list_df8 <- rbindlist(pval.list)
pval.list_df8 <- as.data.frame(pval.list_df8)
# make into long format: where the new column  called "vs" contains the sample sample_type_depth comparison, and new column "p_value" contains the p_values
pval.list_df8 <- gather(pval.list_df8, vs, p_value, 1, factor_key=TRUE)
# adjust p values
pval.list_df8 <- adjust_pvalue(pval.list_df8, p.col = "p_value", output.col = "adj_p", method = "BH")
```

## 8.9. only organic 30-80

```{r}

subset <- subset_samples(ps_RA_nf_genus_pruned, sample_type_depth=="30...40_organic" | sample_type_depth=="40...80_organic")
subset

#create data table
melt_df <-  psmelt(subset)

pval.list <- list()

for (i in all_top20_unique)
{
  # data.frame with the selected taxonomic group
  df <- filter(melt_df, genus==i)
  x <- pairwise.wilcox.test(df$Abundance, df$sample_type_depth,
                            p.adjust.method = "BH")
  x <- as.data.frame(x[["p.value"]])
  x$genus <- i
  x$comparison <- rownames(x)
  pval.list[[i]] <- x
}

pval.list_df9 <- rbindlist(pval.list)
pval.list_df9 <- as.data.frame(pval.list_df9)
# make into long format: where the new column  called "vs" contains the sample sample_type_depth comparison, and new column "p_value" contains the p_values
pval.list_df9 <- gather(pval.list_df9, vs, p_value, 1, factor_key=TRUE)
# adjust p values
pval.list_df9 <- adjust_pvalue(pval.list_df9, p.col = "p_value", output.col = "adj_p", method = "BH")
```

## 8.10. only conventional 0-20

```{r}

subset <- subset_samples(ps_RA_nf_genus_pruned, sample_type_depth=="0...10_conventional" | sample_type_depth=="10...20_conventional")
subset

#create data table
melt_df <-  psmelt(subset)

pval.list <- list()

for (i in all_top20_unique)
{
  # data.frame with the selected taxonomic group
  df <- filter(melt_df, genus==i)
  x <- pairwise.wilcox.test(df$Abundance, df$sample_type_depth,
                            p.adjust.method = "BH")
  x <- as.data.frame(x[["p.value"]])
  x$genus <- i
  x$comparison <- rownames(x)
  pval.list[[i]] <- x
}

pval.list_df10 <- rbindlist(pval.list)
pval.list_df10 <- as.data.frame(pval.list_df10)
# make into long format: where the new column  called "vs" contains the sample sample_type_depth comparison, and new column "p_value" contains the p_values
pval.list_df10 <- gather(pval.list_df10, vs, p_value, 1, factor_key=TRUE)
# adjust p values
pval.list_df10 <- adjust_pvalue(pval.list_df10, p.col = "p_value", output.col = "adj_p", method = "BH")
```


## 8.11. only conventional 10-30

```{r}

subset <- subset_samples(ps_RA_nf_genus_pruned, sample_type_depth=="10...20_conventional" | sample_type_depth=="20...30_conventional")
subset

#create data table
melt_df <-  psmelt(subset)

pval.list <- list()

for (i in all_top20_unique)
{
  # data.frame with the selected taxonomic group
  df <- filter(melt_df, genus==i)
  x <- pairwise.wilcox.test(df$Abundance, df$sample_type_depth,
                            p.adjust.method = "BH")
  x <- as.data.frame(x[["p.value"]])
  x$genus <- i
  x$comparison <- rownames(x)
  pval.list[[i]] <- x
}

pval.list_df11 <- rbindlist(pval.list)
pval.list_df11 <- as.data.frame(pval.list_df11)
# make into long format: where the new column  called "vs" contains the sample sample_type_depth comparison, and new column "p_value" contains the p_values
pval.list_df11 <- gather(pval.list_df11, vs, p_value, 1, factor_key=TRUE)
# adjust p values
pval.list_df11 <- adjust_pvalue(pval.list_df11, p.col = "p_value", output.col = "adj_p", method = "BH")
```


## 8.12. only conventional 20-40

```{r}

subset <- subset_samples(ps_RA_nf_genus_pruned, sample_type_depth=="20...30_conventional" | sample_type_depth=="30...40_conventional")
subset

#create data table
melt_df <-  psmelt(subset)

pval.list <- list()

for (i in all_top20_unique)
{
  # data.frame with the selected taxonomic group
  df <- filter(melt_df, genus==i)
  x <- pairwise.wilcox.test(df$Abundance, df$sample_type_depth,
                            p.adjust.method = "BH")
  x <- as.data.frame(x[["p.value"]])
  x$genus <- i
  x$comparison <- rownames(x)
  pval.list[[i]] <- x
}

pval.list_df12 <- rbindlist(pval.list)
pval.list_df12 <- as.data.frame(pval.list_df12)
# make into long format: where the new column  called "vs" contains the sample sample_type_depth comparison, and new column "p_value" contains the p_values
pval.list_df12 <- gather(pval.list_df12, vs, p_value, 1, factor_key=TRUE)
# adjust p values
pval.list_df12 <- adjust_pvalue(pval.list_df12, p.col = "p_value", output.col = "adj_p", method = "BH")
```

## 8.13. only conventional 30-80

```{r}

subset <- subset_samples(ps_RA_nf_genus_pruned, sample_type_depth=="30...40_conventional" | sample_type_depth=="40...80_conventional")
subset

#create data table
melt_df <-  psmelt(subset)

pval.list <- list()

for (i in all_top20_unique)
{
  # data.frame with the selected taxonomic group
  df <- filter(melt_df, genus==i)
  x <- pairwise.wilcox.test(df$Abundance, df$sample_type_depth,
                            p.adjust.method = "BH")
  x <- as.data.frame(x[["p.value"]])
  x$genus <- i
  x$comparison <- rownames(x)
  pval.list[[i]] <- x
}

pval.list_df13 <- rbindlist(pval.list)
pval.list_df13 <- as.data.frame(pval.list_df13)
# make into long format: where the new column  called "vs" contains the sample sample_type_depth comparison, and new column "p_value" contains the p_values
pval.list_df13 <- gather(pval.list_df13, vs, p_value, 1, factor_key=TRUE)
# adjust p values
pval.list_df13 <- adjust_pvalue(pval.list_df13, p.col = "p_value", output.col = "adj_p", method = "BH")
```

## 8.14. only meadow 0-20

```{r}

subset <- subset_samples(ps_RA_nf_genus_pruned, sample_type_depth=="0...10_meadow" | sample_type_depth=="10...20_meadow")
subset

#create data table
melt_df <-  psmelt(subset)

pval.list <- list()

for (i in all_top20_unique)
{
  # data.frame with the selected taxonomic group
  df <- filter(melt_df, genus==i)
  x <- pairwise.wilcox.test(df$Abundance, df$sample_type_depth,
                            p.adjust.method = "BH")
  x <- as.data.frame(x[["p.value"]])
  x$genus <- i
  x$comparison <- rownames(x)
  pval.list[[i]] <- x
}

pval.list_df14 <- rbindlist(pval.list)
pval.list_df14 <- as.data.frame(pval.list_df14)
# make into long format: where the new column  called "vs" contains the sample sample_type_depth comparison, and new column "p_value" contains the p_values
pval.list_df14 <- gather(pval.list_df14, vs, p_value, 1, factor_key=TRUE)
# adjust p values
pval.list_df14 <- adjust_pvalue(pval.list_df14, p.col = "p_value", output.col = "adj_p", method = "BH")
```

## 8.15. only meadow 10-30

```{r}

subset <- subset_samples(ps_RA_nf_genus_pruned, sample_type_depth=="10...20_meadow" | sample_type_depth=="20...30_meadow")
subset

#create data table
melt_df <-  psmelt(subset)

pval.list <- list()

for (i in all_top20_unique)
{
  # data.frame with the selected taxonomic group
  df <- filter(melt_df, genus==i)
  x <- pairwise.wilcox.test(df$Abundance, df$sample_type_depth,
                            p.adjust.method = "BH")
  x <- as.data.frame(x[["p.value"]])
  x$genus <- i
  x$comparison <- rownames(x)
  pval.list[[i]] <- x
}

pval.list_df15 <- rbindlist(pval.list)
pval.list_df15 <- as.data.frame(pval.list_df15)
# make into long format: where the new column  called "vs" contains the sample sample_type_depth comparison, and new column "p_value" contains the p_values
pval.list_df15 <- gather(pval.list_df15, vs, p_value, 1, factor_key=TRUE)
# adjust p values
pval.list_df15 <- adjust_pvalue(pval.list_df15, p.col = "p_value", output.col = "adj_p", method = "BH")
```

## 8.16. only meadow 20-40

```{r}

subset <- subset_samples(ps_RA_nf_genus_pruned, sample_type_depth=="20...30_meadow" | sample_type_depth=="30...40_meadow")
subset

#create data table
melt_df <-  psmelt(subset)

pval.list <- list()

for (i in all_top20_unique)
{
  # data.frame with the selected taxonomic group
  df <- filter(melt_df, genus==i)
  x <- pairwise.wilcox.test(df$Abundance, df$sample_type_depth,
                            p.adjust.method = "BH")
  x <- as.data.frame(x[["p.value"]])
  x$genus <- i
  x$comparison <- rownames(x)
  pval.list[[i]] <- x
}

pval.list_df16 <- rbindlist(pval.list)
pval.list_df16 <- as.data.frame(pval.list_df16)
# make into long format: where the new column  called "vs" contains the sample sample_type_depth comparison, and new column "p_value" contains the p_values
pval.list_df16 <- gather(pval.list_df16, vs, p_value, 1, factor_key=TRUE)
# adjust p values
pval.list_df16 <- adjust_pvalue(pval.list_df16, p.col = "p_value", output.col = "adj_p", method = "BH")
```

## 8.17. only meadow 30-80

```{r}

subset <- subset_samples(ps_RA_nf_genus_pruned, sample_type_depth=="30...40_meadow" | sample_type_depth=="40...80_meadow")
subset

#create data table
melt_df <-  psmelt(subset)

pval.list <- list()

for (i in all_top20_unique)
{
  # data.frame with the selected taxonomic group
  df <- filter(melt_df, genus==i)
  x <- pairwise.wilcox.test(df$Abundance, df$sample_type_depth,
                            p.adjust.method = "BH")
  x <- as.data.frame(x[["p.value"]])
  x$genus <- i
  x$comparison <- rownames(x)
  pval.list[[i]] <- x
}

pval.list_df17 <- rbindlist(pval.list)
pval.list_df17 <- as.data.frame(pval.list_df17)
# make into long format: where the new column  called "vs" contains the sample sample_type_depth comparison, and new column "p_value" contains the p_values
pval.list_df17 <- gather(pval.list_df17, vs, p_value, 1, factor_key=TRUE)
# adjust p values
pval.list_df17 <- adjust_pvalue(pval.list_df17, p.col = "p_value", output.col = "adj_p", method = "BH")
```




# 11. combine all p-values

```{r}
all.pvals <- rbind(pval.list_df1, pval.list_df2, pval.list_df3, pval.list_df4, pval.list_df5, pval.list_df6, pval.list_df7, pval.list_df8, pval.list_df9, pval.list_df10, pval.list_df11, pval.list_df12, pval.list_df13, pval.list_df14, pval.list_df14, pval.list_df16, pval.list_df17)

all.sig.pvals <- all.pvals[all.pvals$adj_p <= 0.05, ]   

all.sig.genus <- unique(all.sig.pvals$genus)
length(all.sig.genus)

```


# 12. Make ps with the significant taxa and get taxa mean and se values

```{r}
# only keep sig

# filter by taxa name
ps_RA_nf_genus_pruned_sig <- prune_taxa(all.sig.genus, ps_RA_nf_genus_pruned)
ps_Heatmap <- ps_RA_nf_genus_pruned_sig
ps_Heatmap

rm(ps_RA_nf_genus_pruned_sig)

df <- psmelt(ps_Heatmap)

x <- df %>%
  group_by(OTU, sample_type, depth) %>%
  summarise(mean = mean(Abundance, na.rm = TRUE), se = (sd(Abundance, na.rm = TRUE)/sqrt(length((Abundance)))))

x

```


# 13. MAKE HEATMAP


## 13.1. Add FUNGuild annotation for HEATMAP

### 13.1.1. Build FUNGuild


I need to make a separate funguild phyloseq for heatmap, where I have only the genus and higher level annotation (no species level)


```{r}
setwd('C:\\Users\\lehakkin\\PhD\\Fungal_ITS_soil_vertical_profile')

FG <- read.csv2("FUNGuild_31_05_2024.csv")

# what different levels there are:
unique(FG$taxonomicLevel)
```

I will get the annotations from genus and higher tax levels:

Genus

```{r}
fg <- FG[FG$taxonomicLevel == "Genus", ]      
tax_table <- as.data.frame(tax_table(ps))
tax_table$OTU <- rownames(tax_table)

# merge tables
colnames(fg)[1] <- "genus"
FG_tax_table <- merge(tax_table, fg, by = "genus", all.x = TRUE)


# modify "possible"s to NA's at guild and trophic mode -level
FG_tax_table$guild[FG_tax_table$confidenceRanking == "Possible"] <- NA
FG_tax_table$trophicMode[FG_tax_table$confidenceRanking == "Possible"] <- NA

# only keep the OTU, TrophicMode and guild
# and OTU as row names

FG_tax_table <- FG_tax_table[c("OTU", "trophicMode", "guild")]

# and OTU as row names
FG_tax_table <- column_to_rownames(FG_tax_table, var = "OTU")

# change column names
colnames(FG_tax_table)[1] <- "trophicMode_gen"
colnames(FG_tax_table)[2] <- "guild_gen"

# save with new name
FUNGuild_gen <- FG_tax_table


```


Family

```{r}
fg <- FG[FG$taxonomicLevel == "Family", ]      
tax_table <- as.data.frame(tax_table(ps))
tax_table$OTU <- rownames(tax_table)

# merge tables
colnames(fg)[1] <- "family"
FG_tax_table <- merge(tax_table, fg, by = "family", all.x = TRUE)

# modify "possible"s to NA's at guild and trophic mode -level
FG_tax_table$guild[FG_tax_table$confidenceRanking == "Possible"] <- NA
FG_tax_table$trophicMode[FG_tax_table$confidenceRanking == "Possible"] <- NA

# only keep the OTU, TrophicMode and guild
FG_tax_table <- FG_tax_table[c("OTU", "trophicMode", "guild")]

# and taxon as row names
FG_tax_table <- column_to_rownames(FG_tax_table, var = "OTU")

# change column names
colnames(FG_tax_table)[1] <- "trophicMode_fam"
colnames(FG_tax_table)[2] <- "guild_fam"

# save with new name
FUNGuild_fam <- FG_tax_table
```


Order

```{r}
fg <- FG[FG$taxonomicLevel == "Order", ]      
tax_table <- as.data.frame(tax_table(ps))
tax_table$OTU <- rownames(tax_table)

# merge tables
colnames(fg)[1] <- "order"
FG_tax_table <- merge(tax_table, fg, by = "order", all.x = TRUE)

# modify "possible"s to NA's at guild and trophic mode -level
FG_tax_table$guild[FG_tax_table$confidenceRanking == "Possible"] <- NA
FG_tax_table$trophicMode[FG_tax_table$confidenceRanking == "Possible"] <- NA

# only keep the OTU, TrophicMode and guild
FG_tax_table <- FG_tax_table[c("OTU", "trophicMode", "guild")]

# and taxon as row names
FG_tax_table <- column_to_rownames(FG_tax_table, var = "OTU")

# change column names
colnames(FG_tax_table)[1] <- "trophicMode_ord"
colnames(FG_tax_table)[2] <- "guild_ord"

# save with new name
FUNGuild_ord <- FG_tax_table
```


Phylum


```{r}
fg <- FG[FG$taxonomicLevel == "Phylum", ]      
tax_table <- as.data.frame(tax_table(ps))
tax_table$OTU <- rownames(tax_table)

# merge tables
colnames(fg)[1] <- "phylum"
FG_tax_table <- merge(tax_table, fg, by = "phylum", all.x = TRUE)

# modify "possible"s to NA's at guild and trophic mode -level
FG_tax_table$guild[FG_tax_table$confidenceRanking == "Possible"] <- NA
FG_tax_table$trophicMode[FG_tax_table$confidenceRanking == "Possible"] <- NA

# only keep the OTU, TrophicMode and guild
FG_tax_table <- FG_tax_table[c("OTU", "trophicMode", "guild")]

# and OTU as row names
FG_tax_table <- column_to_rownames(FG_tax_table, var = "OTU")

# change column names
colnames(FG_tax_table)[1] <- "trophicMode_phy"
colnames(FG_tax_table)[2] <- "guild_phy"

# save with new name
FUNGuild_phy <- FG_tax_table
```


Combine all annotations:Genus  Family  Order Phylum

```{r}

x <- left_join(rownames_to_column(FUNGuild_gen), rownames_to_column(FUNGuild_fam), by = "rowname")
x <- left_join(x, rownames_to_column(FUNGuild_ord), by = "rowname")
x <- left_join(x, rownames_to_column(FUNGuild_phy), by = "rowname")
```



```{r}

# get the value from another column if NA
y <- x %>%
  mutate(trophicMode_gen = coalesce(trophicMode_gen,trophicMode_fam))

# get the value from another column if NA
y <- y %>%
  mutate(guild_gen = coalesce(guild_gen,guild_fam))

# same for order
# get the value from another column if NA
y <- y %>%
  mutate(trophicMode_gen = coalesce(trophicMode_gen,trophicMode_ord))

# get the value from another column if NA
y <- y %>%
  mutate(guild_gen = coalesce(guild_gen,guild_ord))

# same for phylum
# get the value from another column if NA
y <- y %>%
  mutate(trophicMode_gen = coalesce(trophicMode_gen,trophicMode_phy))

# get the value from another column if NA
y <- y %>%
  mutate(guild_gen = coalesce(guild_gen,guild_phy))

# then rename the gen columns
colnames(y)[2] <- "trophicMode"
colnames(y)[3] <- "guild"


# remove the rest of the columns
y <- y[, -c(4:11)]

# rownames
y2 <- y[,-1]
rownames(y2) <- y[,1]

```


```{r}
# Remove empty spaces and remove also "|"

y2$trophicMode <- gsub(" ", "", y2$trophicMode, fixed = TRUE)
y2$guild <- gsub("|", "", y2$guild, fixed = TRUE)
```



```{r}
y3 <- left_join(rownames_to_column(y2), rownames_to_column(tax_table), by = "rowname")
row.names(y3) <- y3$rowname
y3 <- y3[, -1]

```


Define AMFs, Ectomycorrhizal and Plant pathogens

Here in FUNGuild column:

- Ectomycorrhizal           =   guilds containing "Ectomycorrhizal" from trophic mode Symbiotroph only, NOTE! this is same as pure Ectomycorrhizal!!
- Arbuscular Mycorrhizal    =   all guilds containing "Arbuscular Mycorrhizal" from trophic mode Symbiotroph (no AMF in other trophic modes)
- Endophyte                 =   Pure endophytes from trophic mode Symbiotroph only
- Plant Pathogen            =   Pure Plant Pathogens from trophic mode Pathotroph only


```{r}
z <- y3

z <- z %>%  
   mutate(FG = case_when(grepl("Ectomycorrhizal", guild) & trophicMode=="Symbiotroph" ~ "Ectomycorrhizal", grepl("Arbuscular", guild) ~ "Arbuscular Mycorrhizal", guild == "Endophyte" & trophicMode=="Symbiotroph" ~ "Endophyte", guild=="Plant Pathogen" & trophicMode=="Pathotroph"~ "Plant Pathogen"))
```


```{r}
z <- z %>%
  mutate(FUNGuild = case_when(FG == "Ectomycorrhizal" |  FG == "Arbuscular Mycorrhizal" | FG == "Endophyte" | FG == "Plant Pathogen" ~ FG, guild != "Ectomycorrhizal" |  FG != "Arbuscular Mycorrhizal" |  FG != "Endophyte" | FG != "Plant Pathogen" ~ z$trophicMode))

```



```{r}
FG_tax_table <- z

# remove the FG column

FG_tax_table <- FG_tax_table[, -c(11)]

```


Check the different written forms, if I have empty spaces?

```{r}
unique(FG_tax_table$trophicMode)
#unique(FG_tax_table$guild)
unique(FG_tax_table$FUNGuild)
```



```{r}
# change some names for FUNGuild
FG_tax_table$FUNGuild[FG_tax_table$FUNGuild=="Symbiotroph"] <- "Other Symbiotroph"
FG_tax_table$FUNGuild[FG_tax_table$FUNGuild=="Pathotroph"] <- "Other Pathotroph"
FG_tax_table$FUNGuild[FG_tax_table$FUNGuild=="Pathotroph-Pathotroph-Saprotroph"] <- "Pathotroph-Saprotroph"
FG_tax_table$FUNGuild[FG_tax_table$FUNGuild=="Saprotroph-Saprotroph-Symbiotroph"] <- "Saprotroph-Symbiotroph"

# and for trophicMode
FG_tax_table$trophicMode[FG_tax_table$trophicMode=="Pathotroph-Pathotroph-Saprotroph"] <- "Pathotroph-Saprotroph"
FG_tax_table$trophicMode[FG_tax_table$trophicMode=="Saprotroph-Saprotroph-Symbiotroph"] <- "Saprotroph-Symbiotroph"

# reorder
FG_tax_table <- FG_tax_table[, c(1:2, 11, 3:10)]

```


Check again

```{r}
unique(FG_tax_table$trophicMode)
#unique(FG_tax_table$guild)
unique(FG_tax_table$FUNGuild)
```


### 13.1.2. Make ps_FG for HEATMAP


```{r}

ps_FG_HEATMAP_no_clusters <- phyloseq(otu_table(ps), tax_table(as.matrix(FG_tax_table)), sample_data(ps))
ps_FG_HEATMAP_no_clusters


```


### 13.1.3. Make FIGURE

```{r}
ps_RA <- transform(ps, "compositional")

# Pick relative abundances (compositional) and sample metadata 

ps_RA_nf <- subset_samples(ps_RA, sample_type!="forest")

# You have only submitted samples, not OTUs. Define prevalence of each taxa
# (in how many samples did each taxa appear at least once)
prev0 = apply(X = otu_table(ps_RA_nf),
              MARGIN = ifelse(taxa_are_rows(ps_RA_nf), yes = 1, no = 2),
              FUN = function(x){sum(x > 0)})

# Execute prevalence filter, using `prune_taxa()` function
ps_RA_nf = prune_taxa((prev0 > 0), ps_RA_nf)
ps_RA_nf


all.sig.genus <- unique(all.sig.pvals$genus)
length(all.sig.genus)


# first aggregate to genus
ps_RA_nf_genus <- aggregate_rare(ps_RA_nf, level = "genus", detection = 0/140, prevalence = 0/100)

all_top20_unique <- all_top20_unique_df$all_top20_unique

# filter by taxa name
ps_RA_nf_genus_pruned <- prune_taxa(all_top20_unique, ps_RA_nf_genus)
ps_RA_nf_genus_pruned

# filter by taxa name
ps_RA_nf_genus_pruned_sig <- prune_taxa(all.sig.genus, ps_RA_nf_genus_pruned)
ps_Heatmap <- ps_RA_nf_genus_pruned_sig
ps_Heatmap

```



```{r}
FG_tax <- as.data.frame(as.matrix(tax_table(ps_FG_HEATMAP_no_clusters)))
```


## 13.1.5.Transform RA and z scaled


```{r}
ps_genus <- aggregate_rare(ps, level = 'genus', detection = 0/100, prevalence = 0/100)
ps_genus

# subset
ps_genus_NO_FOREST <- subset_samples(ps_genus, sample_type!="forest")
ps_genus_NO_FOREST

meta_nf <- meta(ps_genus_NO_FOREST)

meta_nf$Depth <- gsub("...", "-", meta_nf$depth, fixed = TRUE)

meta_nf$Depth[meta_nf$Depth=="40-"] <- "40-80"


meta_nf$Depth[meta_nf$Depth=="0-10"] <- "0-10 cm"
meta_nf$Depth[meta_nf$Depth=="10-20"] <- "10-20 cm"
meta_nf$Depth[meta_nf$Depth=="20-30"] <- "20-30 cm"
meta_nf$Depth[meta_nf$Depth=="30-40"] <- "30-40 cm"
meta_nf$Depth[meta_nf$Depth=="40-80"] <- "40-80 cm"

# RA-transform
ps_genus_NO_FOREST_RA <- microbiome::transform(ps_genus_NO_FOREST, 'compositional')

# Z-transform
ps_genus_NO_FOREST_RA_z <- microbiome::transform(ps_genus_NO_FOREST_RA, 'Z', log10 = FALSE)

# subset to only the significant genus from above

hetamap.taxa <- taxa_names(ps_Heatmap)
data <- as(otu_table(ps_genus_NO_FOREST_RA_z), "matrix")
data <- as.data.frame(data)
data_subset <- data[(row.names(data) %in% hetamap.taxa),]

data_subset <- as.matrix(data_subset)

# add annotations "Depth" and "Treatment"
my_sample_col <- data.frame(meta_nf[c("Depth", "sample_type")], row.names = row.names(meta_nf))
colnames(my_sample_col) <- c("Depth", "Treatment")


x <- as.data.frame(t(data_subset))
y <- as.data.frame(my_sample_col)
colnames(y) <- c("Depth", "sample_type")
z <- dplyr::left_join(rownames_to_column(x), rownames_to_column(y), by=c("rowname" = "rowname"))
colnames(z)[1] <- "ID"

f <- z %>%
  group_by(sample_type, Depth) %>%
  summarise_all(mean)

# remove ID column
f <- f[ , -3]

# make new ID column, which is the soiltypedepth
library(stringr)
f$ID <- str_c(f$sample_type, '_', f$Depth)
# make new df with just sample type, depth and sampletypedepth

df <- f[c("sample_type", "Depth", "ID")]
# remove extra columns from f
f2 <- f[, -c(1, 2, 72)]
# column into rownames
rownames(f2) <- f$ID
# make into numeric matrix
f3 <- data.matrix(f2, rownames.force = NA)
f4 <- t(f3)

# same for df
# remove extra columns from df
df2 <- df[ , -c(3)]
# column into rownames
rownames(df2) <- df$ID
my_sample_col2 <- df2
my_sample_col3 <- as.data.frame(my_sample_col2)
rownames(my_sample_col3) <- df$ID
# lets add annotations of samples

# add annotations "Depth" and "soil_type" and change order
colnames(my_sample_col3) <- c("Treatment", "Depth")
my_sample_col3 <- my_sample_col3[, c(2,1)]
```


## 13.1.6 Finally plot HEATMAP

```{r fig.dim = c(10, 16)}

library("pheatmap")
library("ggplotify")

# only keep the genus and FUNGuild
x <- FG_tax[, c(9, 3)]
# Remove duplicates by single column
FUNGuild_tax_table <- x[!duplicated(x$genus), ]
dim(FUNGuild_tax_table)

# and genus as row names
rownames(FUNGuild_tax_table) <- NULL
FUNGuild_tax_table <- column_to_rownames(FUNGuild_tax_table, var = "genus")

# view data frame
unique(FUNGuild_tax_table$FUNGuild)

FUNGuild_tax_table$FUNGuild <- as.factor(FUNGuild_tax_table$FUNGuild)

# change level order
FUNGuild_tax_table$FUNGuild <- factor(FUNGuild_tax_table$FUNGuild, levels = c("Plant Pathogen", "Other Pathotroph", "Pathotroph-Saprotroph", "Pathotroph-Saprotroph-Symbiotroph", "Pathotroph-Symbiotroph", "Saprotroph", "Saprotroph-Symbiotroph", "Other Symbiotroph", "Ectomycorrhizal", "Endophyte", "Arbuscular Mycorrhizal"))
levels(FUNGuild_tax_table$FUNGuild)


#Create color palette

my_colour = list(
  "Treatment" = c(meadow = "#fbc02d", organic = "#8a8a8a", conventional =  "#b71c1c"),
  Depth = c('0-10 cm' = "#387212", '10-20 cm' = "#ADC476", '20-30 cm' = "#D8D2BA",'30-40 cm' = "#907852", '40-80 cm' = "#6A4C3A"),
  FUNGuild = c("Plant Pathogen" = "deeppink", "Other Pathotroph" = "#d6849a", "Pathotroph-Saprotroph" = "#e3adbc", "Pathotroph-Saprotroph-Symbiotroph" = "#f1d6dd", "Pathotroph-Symbiotroph" = "#faf1f4", "Saprotroph" = "#CBBEAD", "Saprotroph-Symbiotroph" = "darkseagreen", "Other Symbiotroph" = "lightgreen", "Ectomycorrhizal" = "darkgreen", "Endophyte" = "#A2CF31", "Arbuscular Mycorrhizal" = "darkolivegreen1"))

# add gap between sample_types
gp_col = which(diff(as.numeric(factor(my_sample_col3$Treatment)))!=0)

p2 <- as.ggplot(function() pheatmap(f4, cluster_cols = FALSE, cluster_rows = TRUE, annotation_col = my_sample_col3, annotation_colors = my_colour, gaps_col = gp_col, color=colorRampPalette(c("navy", "white", "red"))(50), show_colnames = FALSE, legend = TRUE, annotation_row = FUNGuild_tax_table, border_color = NA, cellheight = 16, fontsize = 14, fontsize_row = 14, annotation_names_row = FALSE, annotation_names_col = FALSE))

p2
```



# B) HEATMAP for FOREST ONLY fungal genera no filtering by significant genera


## 1. load packages and data

```{r}

library('phyloseq')
library("cowplot")
library("dplyr")
library("ggplot2")
library("vegan")
library("microbiome")
library("tibble")


setwd('C:\\Users\\lehakkin\\PhD\\Fungal_ITS_soil_vertical_profile')

load('ps_FINAL')
ps


# Pick relative abundances (compositional) and sample metadata 
ps_RA <- microbiome::transform(ps, "compositional")
ps_RA_nf <- subset_samples(ps_RA, sample_type=="forest")

# You have only submitted samples, not OTUs. Define prevalence of each taxa
# (in how many samples did each taxa appear at least once)
prev0 = apply(X = otu_table(ps_RA_nf),
              MARGIN = ifelse(taxa_are_rows(ps_RA_nf), yes = 1, no = 2),
              FUN = function(x){sum(x > 0)})

# Execute prevalence filter, using `prune_taxa()` function
ps_RA_nf = prune_taxa((prev0 > 0), ps_RA_nf)
ps_RA_nf

meta_nf <- meta(ps_RA_nf)
```


# 2. Overview of the heatmap analysis

I will use RA values
For the final heatmap figure the RA values will be standardizes (z-transformed)

1) First pick the **10** most abundant genera in each soil type depth
2) Make a HEATMAP of the most abundant taxa only


# 3. at genus level

```{r}

ps_genus <- aggregate_rare(ps, level = 'genus', detection = 0/100, prevalence = 0/140)
ps_genus


ps_genus_RA <- microbiome::transform(ps_genus, "compositional")

ps_genus_RA_pruned <- subset_samples(ps_genus_RA, sample_type=="forest")
ps_genus_RA_pruned
```


# 4. Pick 10 most abund taxa in each forest soil layer

```{r}
# sample wise filtering according to most abund. genera

# Initialize an empty list to store the taxa
abund.taxa <- list()

for (i in meta_nf$sample_type) {
  for (j in meta_nf$depth) {
    x <- sample_names(sample_data(ps_genus_RA_pruned)[sample_data(ps_genus_RA_pruned)$sample_type == i & sample_data(ps_genus_RA_pruned)$depth == j,])

# Calculate taxa mean of the selected samples
top10 <- head(sort(rowMeans(otu_table(ps_genus_RA_pruned)[,x]), decreasing = TRUE), 10)

result_name <- paste("sample_type", i, "depth", j, sep = "_")

abund.taxa[[result_name]] <- top10
  }
}

```



# 5. Combine picked taxa

```{r}
management_layer <- c(names(abund.taxa))

all_top10 <- c()

for (i in management_layer) {
  top10 <- c(names(abund.taxa[[i]]))
  all_top10 <- c(all_top10, top10)

}

all_top10_unique <- unique(all_top10)
length(all_top10_unique)
```


# 6. Make a phyloseq of picked taxa

```{r}
# first aggregate to genus
ps_RA_nf_genus <- aggregate_rare(ps_RA_nf, level = "genus", detection = 0/140, prevalence = 0/100)

# filter by taxa name
ps_RA_nf_genus_pruned <- prune_taxa(all_top10_unique, ps_RA_nf_genus)
ps_RA_nf_genus_pruned

setwd('C:\\Users\\lehakkin\\PhD\\Fungal_ITS_soil_vertical_profile')

ps_genus_FOREST_ONLY_HETAMAP <- ps_RA_nf_genus_pruned
#save(ps_genus_FOREST_ONLY_HETAMAP, file='ps_genus_nf_HETAMAP_all_top10_FOREST_ONLY_unclassified_included')
```


# 9. Get taxa mean and se values

```{r}
ps_Heatmap <- ps_genus_FOREST_ONLY_HETAMAP
ps_Heatmap

df <- psmelt(ps_Heatmap)

x <- df %>%
  group_by(OTU, sample_type, depth) %>%
  summarise(mean = mean(Abundance, na.rm = TRUE), se = (sd(Abundance, na.rm = TRUE)/sqrt(length((Abundance)))))

x

```


# 13. MAKE HEATMAP


## 13.1. Add FUNGuild annotation for HEATMAP

### 13.1.1. Build FUNGuild


I need to make a separate funguild phyloseq for heatmap, where I have only the genus and higher level annotation (no species level)


```{r}
setwd('C:\\Users\\lehakkin\\PhD\\Fungal_ITS_soil_vertical_profile')

FG <- read.csv2("FUNGuild_31_05_2024.csv")

# what different levels there are:
unique(FG$taxonomicLevel)
```

I will get the annotations from genus and higher tax levels:

Genus

```{r}
fg <- FG[FG$taxonomicLevel == "Genus", ]      
tax_table <- as.data.frame(tax_table(ps))
tax_table$OTU <- rownames(tax_table)

# merge tables
colnames(fg)[1] <- "genus"
FG_tax_table <- merge(tax_table, fg, by = "genus", all.x = TRUE)


# modify "possible"s to NA's at guild and trophic mode -level
FG_tax_table$guild[FG_tax_table$confidenceRanking == "Possible"] <- NA
FG_tax_table$trophicMode[FG_tax_table$confidenceRanking == "Possible"] <- NA

# only keep the OTU, TrophicMode and guild
# and OTU as row names

FG_tax_table <- FG_tax_table[c("OTU", "trophicMode", "guild")]

# and OTU as row names
FG_tax_table <- column_to_rownames(FG_tax_table, var = "OTU")

# change column names
colnames(FG_tax_table)[1] <- "trophicMode_gen"
colnames(FG_tax_table)[2] <- "guild_gen"

# save with new name
FUNGuild_gen <- FG_tax_table


```


Family

```{r}
fg <- FG[FG$taxonomicLevel == "Family", ]      
tax_table <- as.data.frame(tax_table(ps))
tax_table$OTU <- rownames(tax_table)

# merge tables
colnames(fg)[1] <- "family"
FG_tax_table <- merge(tax_table, fg, by = "family", all.x = TRUE)

# modify "possible"s to NA's at guild and trophic mode -level
FG_tax_table$guild[FG_tax_table$confidenceRanking == "Possible"] <- NA
FG_tax_table$trophicMode[FG_tax_table$confidenceRanking == "Possible"] <- NA

# only keep the OTU, TrophicMode and guild
FG_tax_table <- FG_tax_table[c("OTU", "trophicMode", "guild")]

# and taxon as row names
FG_tax_table <- column_to_rownames(FG_tax_table, var = "OTU")

# change column names
colnames(FG_tax_table)[1] <- "trophicMode_fam"
colnames(FG_tax_table)[2] <- "guild_fam"

# save with new name
FUNGuild_fam <- FG_tax_table
```


Order

```{r}
fg <- FG[FG$taxonomicLevel == "Order", ]      
tax_table <- as.data.frame(tax_table(ps))
tax_table$OTU <- rownames(tax_table)

# merge tables
colnames(fg)[1] <- "order"
FG_tax_table <- merge(tax_table, fg, by = "order", all.x = TRUE)

# modify "possible"s to NA's at guild and trophic mode -level
FG_tax_table$guild[FG_tax_table$confidenceRanking == "Possible"] <- NA
FG_tax_table$trophicMode[FG_tax_table$confidenceRanking == "Possible"] <- NA

# only keep the OTU, TrophicMode and guild
FG_tax_table <- FG_tax_table[c("OTU", "trophicMode", "guild")]

# and taxon as row names
FG_tax_table <- column_to_rownames(FG_tax_table, var = "OTU")

# change column names
colnames(FG_tax_table)[1] <- "trophicMode_ord"
colnames(FG_tax_table)[2] <- "guild_ord"

# save with new name
FUNGuild_ord <- FG_tax_table
```


Phylum


```{r}
fg <- FG[FG$taxonomicLevel == "Phylum", ]      
tax_table <- as.data.frame(tax_table(ps))
tax_table$OTU <- rownames(tax_table)

# merge tables
colnames(fg)[1] <- "phylum"
FG_tax_table <- merge(tax_table, fg, by = "phylum", all.x = TRUE)

# modify "possible"s to NA's at guild and trophic mode -level
FG_tax_table$guild[FG_tax_table$confidenceRanking == "Possible"] <- NA
FG_tax_table$trophicMode[FG_tax_table$confidenceRanking == "Possible"] <- NA

# only keep the OTU, TrophicMode and guild
FG_tax_table <- FG_tax_table[c("OTU", "trophicMode", "guild")]

# and OTU as row names
FG_tax_table <- column_to_rownames(FG_tax_table, var = "OTU")

# change column names
colnames(FG_tax_table)[1] <- "trophicMode_phy"
colnames(FG_tax_table)[2] <- "guild_phy"

# save with new name
FUNGuild_phy <- FG_tax_table
```


Combine all annotations:Genus  Family  Order Phylum

```{r}

x <- left_join(rownames_to_column(FUNGuild_gen), rownames_to_column(FUNGuild_fam), by = "rowname")
x <- left_join(x, rownames_to_column(FUNGuild_ord), by = "rowname")
x <- left_join(x, rownames_to_column(FUNGuild_phy), by = "rowname")
```



```{r}

# get the value from another column if NA
y <- x %>%
  mutate(trophicMode_gen = coalesce(trophicMode_gen,trophicMode_fam))

# get the value from another column if NA
y <- y %>%
  mutate(guild_gen = coalesce(guild_gen,guild_fam))

# same for order
# get the value from another column if NA
y <- y %>%
  mutate(trophicMode_gen = coalesce(trophicMode_gen,trophicMode_ord))

# get the value from another column if NA
y <- y %>%
  mutate(guild_gen = coalesce(guild_gen,guild_ord))

# same for phylum
# get the value from another column if NA
y <- y %>%
  mutate(trophicMode_gen = coalesce(trophicMode_gen,trophicMode_phy))

# get the value from another column if NA
y <- y %>%
  mutate(guild_gen = coalesce(guild_gen,guild_phy))

# then rename the gen columns
colnames(y)[2] <- "trophicMode"
colnames(y)[3] <- "guild"


# remove the rest of the columns
y <- y[, -c(4:11)]

# rownames
y2 <- y[,-1]
rownames(y2) <- y[,1]

```


```{r}
# Remove empty spaces and remove also "|"

y2$trophicMode <- gsub(" ", "", y2$trophicMode, fixed = TRUE)
y2$guild <- gsub("|", "", y2$guild, fixed = TRUE)
```



```{r}
y3 <- left_join(rownames_to_column(y2), rownames_to_column(tax_table), by = "rowname")
row.names(y3) <- y3$rowname
y3 <- y3[, -1]

```


Define AMFs, Ectomycorrhizal and Plant pathogens

Here in FUNGuild column:

- Ectomycorrhizal           =   guilds containing "Ectomycorrhizal" from trophic mode Symbiotroph only, NOTE! this is same as pure Ectomycorrhizal!!
- Arbuscular Mycorrhizal    =   all guilds containing "Arbuscular Mycorrhizal" from trophic mode Symbiotroph (no AMF in other trophic modes)
- Endophyte                 =   Pure endophytes from trophic mode Symbiotroph only
- Plant Pathogen            =   Pure Plant Pathogens from trophic mode Pathotroph only


```{r}
z <- y3

z <- z %>%  
   mutate(FG = case_when(grepl("Ectomycorrhizal", guild) & trophicMode=="Symbiotroph" ~ "Ectomycorrhizal", grepl("Arbuscular", guild) ~ "Arbuscular Mycorrhizal", guild == "Endophyte" & trophicMode=="Symbiotroph" ~ "Endophyte", guild=="Plant Pathogen" & trophicMode=="Pathotroph"~ "Plant Pathogen"))
```


```{r}
z <- z %>%
  mutate(FUNGuild = case_when(FG == "Ectomycorrhizal" |  FG == "Arbuscular Mycorrhizal" | FG == "Endophyte" | FG == "Plant Pathogen" ~ FG, guild != "Ectomycorrhizal" |  FG != "Arbuscular Mycorrhizal" |  FG != "Endophyte" | FG != "Plant Pathogen" ~ z$trophicMode))

```



```{r}
FG_tax_table <- z

# remove the FG column

FG_tax_table <- FG_tax_table[, -c(11)]

```


Check the different written forms, if I have empty spaces?

```{r}
unique(FG_tax_table$trophicMode)
#unique(FG_tax_table$guild)
unique(FG_tax_table$FUNGuild)
```



```{r}
# change some names for FUNGuild
FG_tax_table$FUNGuild[FG_tax_table$FUNGuild=="Symbiotroph"] <- "Other Symbiotroph"
FG_tax_table$FUNGuild[FG_tax_table$FUNGuild=="Pathotroph"] <- "Other Pathotroph"
FG_tax_table$FUNGuild[FG_tax_table$FUNGuild=="Pathotroph-Pathotroph-Saprotroph"] <- "Pathotroph-Saprotroph"
FG_tax_table$FUNGuild[FG_tax_table$FUNGuild=="Saprotroph-Saprotroph-Symbiotroph"] <- "Saprotroph-Symbiotroph"

# and for trophicMode
FG_tax_table$trophicMode[FG_tax_table$trophicMode=="Pathotroph-Pathotroph-Saprotroph"] <- "Pathotroph-Saprotroph"
FG_tax_table$trophicMode[FG_tax_table$trophicMode=="Saprotroph-Saprotroph-Symbiotroph"] <- "Saprotroph-Symbiotroph"

# reorder
FG_tax_table <- FG_tax_table[, c(1:2, 11, 3:10)]

```


Check again

```{r}
unique(FG_tax_table$trophicMode)
#unique(FG_tax_table$guild)
unique(FG_tax_table$FUNGuild)
```


### 13.1.2. Save ps_FG for HEATMAP


```{r}
ps_FG_HEATMAP_no_clusters <- phyloseq(otu_table(ps), tax_table(as.matrix(FG_tax_table)), sample_data(ps))
ps_FG_HEATMAP_no_clusters

```


### 13.1.3.


```{r}
FG_tax <- as.data.frame(as.matrix(tax_table(ps_FG_HEATMAP_no_clusters)))
```


## 13.1.5.Transform RA and z scaled

```{r}
ps_genus <- aggregate_rare(ps, level = 'genus', detection = 0/100, prevalence = 0/100)
ps_genus

# subset
ps_genus_NO_FOREST <- subset_samples(ps_genus, sample_type=="forest")
ps_genus_NO_FOREST

meta_nf <- meta(ps_genus_NO_FOREST)

meta_nf$Depth <- gsub("...", "-", meta_nf$depth, fixed = TRUE)

meta_nf$Depth[meta_nf$Depth=="40-"] <- "40-80"


meta_nf$Depth[meta_nf$Depth=="0-10"] <- "0-10 cm"
meta_nf$Depth[meta_nf$Depth=="10-20"] <- "10-20 cm"
meta_nf$Depth[meta_nf$Depth=="20-30"] <- "20-30 cm"
meta_nf$Depth[meta_nf$Depth=="30-40"] <- "30-40 cm"
meta_nf$Depth[meta_nf$Depth=="40-80"] <- "40-80 cm"

# RA-transform
ps_genus_NO_FOREST_RA <- microbiome::transform(ps_genus_NO_FOREST, 'compositional')

# Z-transform
ps_genus_NO_FOREST_RA_z <- microbiome::transform(ps_genus_NO_FOREST_RA, 'Z', log10 = FALSE)

# subset to only the 30 genera from above

hetamap.taxa <- taxa_names(ps_Heatmap)
data <- as(otu_table(ps_genus_NO_FOREST_RA_z), "matrix")
data <- as.data.frame(data)
data_subset <- data[(row.names(data) %in% hetamap.taxa),]
# now only 30 genuses, as should be
data_subset <- as.matrix(data_subset)

# add annotations "Depth" and "Treatment"
my_sample_col <- data.frame(meta_nf[c("Depth", "sample_type")], row.names = row.names(meta_nf))
colnames(my_sample_col) <- c("Depth", "Treatment")


x <- as.data.frame(t(data_subset))
y <- as.data.frame(my_sample_col)
colnames(y) <- c("Depth", "sample_type")
z <- dplyr::left_join(rownames_to_column(x), rownames_to_column(y), by=c("rowname" = "rowname"))
colnames(z)[1] <- "ID"

f <- z %>%
  group_by(sample_type, Depth) %>%
  summarise_all(mean)

# remove ID column
f <- f[ , -3]

# make new ID column, which is the soiltypedepth
library(stringr)
f$ID <- str_c(f$sample_type, '_', f$Depth)
# make new df with just sample type, Depth and sampletypedepth

df <- f[c("sample_type", "Depth", "ID")]
# remove extra columns from f
f2 <- f[, -c(1, 2, 34)]
# column into rownames
rownames(f2) <- f$ID
# make into numeric matrix
f3 <- data.matrix(f2, rownames.force = NA)
f4 <- t(f3)

# same for df
# remove extra columns from df
df2 <- df[ , -c(3)]
# column into rownames
rownames(df2) <- df$ID
my_sample_col2 <- df2
my_sample_col3 <- as.data.frame(my_sample_col2)
rownames(my_sample_col3) <- df$ID
# lets add annotations of samples

# add annotations "Depth" and "soil_type" and change order
colnames(my_sample_col3) <- c("Treatment", "Depth")
my_sample_col3 <- my_sample_col3[, c(2,1)]
```


## 13.1.6 Finally plot HEATMAP

```{r fig.dim = c(10, 12)}
library("pheatmap")

# only keep the genus and FUNGuild
x <- FG_tax[, c(9, 3)]
# Remove duplicates by single column
FUNGuild_tax_table <- x[!duplicated(x$genus), ]
dim(FUNGuild_tax_table)

# and genus as row names
rownames(FUNGuild_tax_table) <- NULL
FUNGuild_tax_table <- column_to_rownames(FUNGuild_tax_table, var = "genus")

# view data frame
unique(FUNGuild_tax_table$FUNGuild)

FUNGuild_tax_table$FUNGuild <- as.factor(FUNGuild_tax_table$FUNGuild)

# change level order
FUNGuild_tax_table$FUNGuild <- factor(FUNGuild_tax_table$FUNGuild, levels = c("Plant Pathogen", "Other Pathotroph", "Pathotroph-Saprotroph", "Pathotroph-Saprotroph-Symbiotroph", "Pathotroph-Symbiotroph", "Saprotroph", "Saprotroph-Symbiotroph", "Other Symbiotroph", "Ectomycorrhizal", "Endophyte", "Arbuscular Mycorrhizal"))
levels(FUNGuild_tax_table$FUNGuild)


#Create color palette

my_colour = list(
  "Treatment" = c(forest = "#1167b1"),
  Depth = c('0-10 cm' = "#387212", '10-20 cm' = "#ADC476", '20-30 cm' = "#D8D2BA",'30-40 cm' = "#907852", '40-80 cm' = "#6A4C3A"),
  FUNGuild = c("Plant Pathogen" = "deeppink", "Other Pathotroph" = "#d6849a", "Pathotroph-Saprotroph" = "#e3adbc", "Pathotroph-Saprotroph-Symbiotroph" = "#f1d6dd", "Pathotroph-Symbiotroph" = "#faf1f4", "Saprotroph" = "#CBBEAD", "Saprotroph-Symbiotroph" = "darkseagreen", "Other Symbiotroph" = "lightgreen", "Ectomycorrhizal" = "darkgreen", "Endophyte" = "#A2CF31", "Arbuscular Mycorrhizal" = "darkolivegreen1"))

p3 <- as.ggplot(function() pheatmap(f4, cluster_cols = FALSE, cluster_rows = TRUE, annotation_col = my_sample_col3, annotation_colors = my_colour, color=colorRampPalette(c("navy", "white", "red"))(50), show_colnames = FALSE, legend = TRUE, annotation_row = FUNGuild_tax_table, border_color = NA, cellheight = 16, fontsize = 14, fontsize_row = 14, annotation_names_row = FALSE, annotation_names_col = FALSE))

p3
```


# C) COMPOSTION PLOTS


```{r}

library('phyloseq')
library("cowplot")
library("dplyr")
library("ggplot2")
library("vegan")
library("microbiome")
library("tibble")
library(stringr)
library(reshape2)
library(tidyr)

library(metagMisc)
library(pheatmap)
library(metagMisc)
library(RColorBrewer)
library(viridis)
library(tidyverse)
library(ggpubr)

setwd('C:\\Users\\lehakkin\\PhD\\Fungal_ITS_soil_vertical_profile')

load('ps_FINAL')
ps


# Pick relative abundances (compositional) and sample metadata 
ps_RA <- microbiome::transform(ps, "compositional")

load(file = 'ps_FG_with_NAs')#ps_FG
ps_FG


```


# 1. FUNGuild composition

Note, due to reviewers comment, I will redo the Functional guild composition plot so that Unassigned are kept!!!

```{r}
FG_tax <- ps_FG %>% tax_table() %>% as.data.frame()

unique(FG_tax$trophicMode)


```

## 1.1. separate pure Plant pathogens

Note! In the composition figure:

- Ectomycorrhizal           =   guilds containing "Ectomycorrhizal" from trophic mode Symbiotroph only, NOTE! thiis is same as pure Ectomycorrhizal!!
- Arbuscular Mycorrhizal    =   all guilds containing "Arbuscular Mycorrhizal" from trophic mode Symbiotroph (nor AMF in other TMs)
- Endophyte                 =   Pure endophytes from trophic mode Symbiotroph only
- Plant Pathogen            =   Pure Plant Pathogens from trophic mode Pathotroph only


```{r}
z <- FG_tax

z <- z %>%  
   mutate(FG = case_when(grepl("Ectomycorrhizal", guild) & trophicMode=="Symbiotroph" ~ "Ectomycorrhizal", grepl("Arbuscular", guild) ~ "Arbuscular Mycorrhizal", guild == "Endophyte" & trophicMode=="Symbiotroph" ~ "Endophyte", guild=="Plant Pathogen" & trophicMode=="Pathotroph"~ "Plant Pathogen"))
```


```{r}
z <- z %>%
  mutate(FUNGuild = case_when(FG == "Ectomycorrhizal" |  FG == "Arbuscular Mycorrhizal" | FG == "Endophyte" | FG == "Plant Pathogen" ~ FG, guild != "Ectomycorrhizal" |  FG != "Arbuscular Mycorrhizal" |  FG != "Endophyte" | FG != "Plant Pathogen" ~ z$trophicMode))

```


```{r}
# change some names for FUNGuild
z$FUNGuild[z$FUNGuild=="Symbiotroph"] <- "Other Symbiotroph"
z$FUNGuild[z$FUNGuild=="Pathotroph"] <- "Other Pathotroph"

```


```{r}

# remove species and FG
tax <- z[, -c(10, 12)]

unique(tax$FUNGuild)

# rename FUNGuild to species
colnames(tax)[3] <- "species"
tax <- tax %>% as.matrix()

# reassign to phyloseq
x <- ps_FG
tax_table(x) <- tax_table(tax)

x <- aggregate_rare(x, level = "species", detection = 0, prevalence = 0)
x               
# 12 taxa and 140 samples

# lets not remove NAs!!

# remove "Unknown"
#allTaxa = taxa_names(x)
#badTaxa = c("Unknown")
#myTaxa <- allTaxa[!(allTaxa %in% badTaxa)]
#x <- prune_taxa(myTaxa, x)
#x
# would be 11 taxa and 140 samples

x_RA <- transform(x, 'compositional')
```


## 1.2. Plot

```{r fig.dim = c(10, 8)}
#create data table
df <-  psmelt(x_RA)

sampletype_names <- list(
  'forest' = "forest", 'meadow' = "meadow", 'organic' = "organic", 'conventional' = "conventional"
)

sampletype_labeller <- function(variable,value){
  return(sampletype_names[value])
}

df$species <- factor(df$species)

levels(df$species)


# change level order

df$species <- factor(df$species, levels = c("Unknown","Plant Pathogen", "Other Pathotroph", "Pathotroph-Saprotroph", "Pathotroph-Saprotroph-Symbiotroph", "Pathotroph-Symbiotroph", "Saprotroph", "Saprotroph-Symbiotroph", "Other Symbiotroph", "Ectomycorrhizal", "Endophyte", "Arbuscular Mycorrhizal"))
levels(df$species)


#Create color palette

my_colour = list(
  "sample" = c(meadow = "#fbc02d", organic = "#8a8a8a", conventional =  "#b71c1c"),
  depth = c('0...10' = "#387212", '10...20' = "#ADC476", '20...30' = "#D8D2BA",'30...40' = "#907852", '40...' = "#6A4C3A"),
  FUNGuild = c("Unknown" = "grey", "Plant Pathogen" = "deeppink", "Other Pathotroph" = "#d6849a", "Pathotroph-Saprotroph" = "#e3adbc", "Pathotroph-Saprotroph-Symbiotroph" = "#f1d6dd", "Pathotroph-Symbiotroph" = "#faf1f4", "Saprotroph" = "#CBBEAD", "Saprotroph-Symbiotroph" = "darkseagreen", "Other Symbiotroph" = "lightgreen", "Ectomycorrhizal" = "darkgreen", "Endophyte" = "#A2CF31", "Arbuscular Mycorrhizal" = "darkolivegreen1"))


# Make new depth variable

df$Depth <- df$depth

df$Depth <- gsub("...", "-", df$Depth, fixed = TRUE)

df$Depth[df$Depth=="40-"] <- "40-80"


df$Depth[df$Depth=="0-10"] <- "0-10 cm"
df$Depth[df$Depth=="10-20"] <- "10-20 cm"
df$Depth[df$Depth=="20-30"] <- "20-30 cm"
df$Depth[df$Depth=="30-40"] <- "30-40 cm"
df$Depth[df$Depth=="40-80"] <- "40-80 cm"

FG  <- ggplot(df, aes(x = Depth ,y = Abundance, fill = species)) + geom_bar(stat="identity", position="fill") + scale_fill_manual(values = my_colour[["FUNGuild"]]) + facet_grid (cols = vars(sample_type), labeller=sampletype_labeller) + geom_bar(stat="identity", position="fill") + theme(panel.border = element_rect(colour = "black", fill=NA, size=0.5)) + theme_cowplot() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=18),
        legend.text = element_text(size=16),
        legend.title = element_text(size=18),
        legend.spacing.y = unit(0, 'cm'),
        legend.key.size = unit(0.8, 'cm'),
        title = element_text(size=18))+ theme(axis.text.x = element_text(angle = 45, hjust = 1)) + guides(fill = guide_legend(nrow = 4, title="")) + theme(strip.text.x = element_text(size = 22)) +
  ylab(label = "Relative abundance") + theme(legend.position="top") +  xlab("Depth")

FG

```



# 2. Class composition

```{r fig.dim = c(10, 10)}
ps_RA <- microbiome::transform(ps, 'compositional')

ps_RA.class <- aggregate_rare(ps_RA, level = 'class', detection = 3/100, prevalence = 3/140, include.lowest = TRUE)
ps_RA.class

#create data table
ps_RA.class_df <-  psmelt(ps_RA.class)

#Create color palette
#set color palette from RColorBrewer
# Define the number of colors you want
library("RColorBrewer") # nice color options
nb.cols = length(unique((as.data.frame(ps_RA.class@tax_table))$class))
cbbPalette <- colorRampPalette(brewer.pal(12, "Set3"))(nb.cols)

sampletype_names <- list(
  'forest' = "forest", 'meadow' = "meadow", 'organic' = "organic", 'conventional' = "conventional"
)

sampletype_labeller <- function(variable,value){
  return(sampletype_names[value])
}

# check unique values for class
unique(ps_RA.class_df$class)

is.factor(ps_RA.class_df$class)
ps_RA.class_df$class <- as.factor(ps_RA.class_df$class)
levels(ps_RA.class_df$class)

ps_RA.class_df$class <- relevel(ps_RA.class_df$class, "Other")

cbbPalette <- c("#b2b2b2", "#8DD3C7", "#FFED6F", "#CAAEC5", "#F68378", "#8D6942", "#F3B962", "#BCD868", "#6E99BE", "#F0D1E1", "#C191C2", "#FFFFC6", "darkgreen", "#D0D9CD", "#8BC081", "#FF8DB5")



# plot with detection = 1/100, prevalence = 2/100
#Create a plot
classF  <- ggplot(ps_RA.class_df, aes(x = depth ,y = Abundance, fill = class)) + geom_bar(stat="identity", position="fill") + scale_fill_manual(values = cbbPalette) + facet_grid (cols = vars(sample_type), labeller=sampletype_labeller) + geom_bar(stat="identity", position="fill") + scale_fill_manual(values = cbbPalette) + theme(panel.border = element_rect(colour = "black", fill=NA, size=0.5)) + theme_cowplot() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=18),
        legend.text = element_text(size=16),
        legend.title = element_text(size=18),
        legend.spacing.y = unit(0, 'cm'),
        legend.key.size = unit(0.8, 'cm'),
        title = element_text(size=18))+ theme(axis.text.x = element_text(angle = 45, hjust = 1)) + guides(fill = guide_legend(nrow = 6, title="")) + theme(strip.text.x = element_text(size = 22)) +
  ylab(label = "Relative abundance") + theme(legend.position="top") +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.x = element_blank())

classF
```


# 3. Phylum composition


```{r fig.dim = c(10, 5)}

ps_RA_phyla_aggr <- aggregate_rare(ps_RA, level = 'phylum', detection = 2/100, prevalence = 2/140)
ps_RA_phyla_aggr

#create data table
ps_RA_phyla_df <-  psmelt(ps_RA_phyla_aggr)

#Create color palette
cbbPalette <- c("#666666","#1B9E77", "#D95F02", "#E7298A", "#7570B3", "#66A61E")

sampletype_names <- list(
  'forest' = "forest", 'meadow' = "meadow", 'organic' = "organic", 'conventional' = "conventional"
)

sampletype_labeller <- function(variable,value){
  return(sampletype_names[value])
}


# check unique values for phylum
unique(ps_RA_phyla_df$phylum)
# [1] "Ascomycota"        "Basidiomycota"     "Glomeromycota"     "Mortierellomycota"
# [5] "Rozellomycota"     "Other"  

ps_RA_phyla_df$phylum <- as.factor(ps_RA_phyla_df$phylum)
levels(ps_RA_phyla_df$phylum)
ps_RA_phyla_df$phylum <- factor(ps_RA_phyla_df$phylum, levels = c("Other","Ascomycota", "Basidiomycota", "Glomeromycota", "Mortierellomycota", "Rozellomycota"))
levels(ps_RA_phyla_df$phylum)

#Create a plot
phylumF  <- ggplot(ps_RA_phyla_df, aes(x = depth ,y = Abundance, fill = phylum)) + geom_bar(stat="identity", position="fill") + scale_fill_manual(values = cbbPalette) + facet_grid (cols = vars(sample_type), labeller=sampletype_labeller) + geom_bar(stat="identity", position="fill") + scale_fill_manual(values = cbbPalette) + theme(panel.border = element_rect(colour = "black", fill=NA, size=0.5)) + theme_cowplot() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=18),
        legend.text = element_text(size=16),
        legend.title = element_text(size=18),
        legend.spacing.y = unit(0, 'cm'),
        legend.key.size = unit(0.8, 'cm'),
        title = element_text(size=18))+ theme(axis.text.x = element_text(angle = 45, hjust = 1)) + guides(fill = guide_legend(title="")) + theme(strip.text.x = element_text(size = 22)) +
  ylab(label = "Relative abundance") + theme(legend.position="top") +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.x = element_blank())

phylumF
```


# D) Combine composition plots and heatmap figures

## 1.1. Composition bar plots

```{r fig.dim = c(10, 24)}
library(ggplotify)


left <- ggarrange(phylumF, classF , FG,
                  labels = c("A", "B", "C"),
                  ncol = 1, nrow = 3, heights = c(1.1, 1.5, 1.5))

left
```



## 2. Combine Heatmaps

```{r fig.dim = c(10, 24)}
right <- ggarrange(p2, p3,
                       labels = c("D", "E"),
                       ncol = 1, nrow = 2, heights = c(2.7, 1.4))

right

```


## 3. Combine composition and heatmap

```{r fig.dim = c(18, 24)}
figure <- ggarrange(left, right,
                    ncol = 2, nrow = 1, heights = c(1, 1), widths = c(1, 1))

figure

```

Saved with width 2200 and height 2600

